---
title: "gene_metadata.json"
output: html_notebook
---

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```

# Validation run metadata
```{r echo=FALSE}
# add file-specific values here
name <- "gene_metadata"
old_synId <-  "syn27644951" 
new_synId <-  "syn27572743"

# not file-specific
old_name <- paste(name, "old", sep="_")
new_name <- paste(name, "new", sep="_")
rules <- paste0("rules_", name)

# print validation summary
cat( name, ".json", "\n", date(),
    "\nold: ", old_synId, " \nnew: ", new_synId,
    sep="")
```

# Downloaded files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
download_files(old_synId, new_synId, name)
```

# Compared old & new files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
compare(get(old_name), get(new_name), name)
```


# Validated new file against rules
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_gene_metadata.R")
test(get(new_name), get(rules))
```


***
# Validation failure investigations
***

## !is.na(hgnc_symbol) / field_length(hgnc_symbol, min = 0, max = 100): 15,996 fails
## !is.na(name) / FALSE	field_length(name, min = 0, max = 200): 15,996 fails (same as above)
## !is.na(name) / FALSE	field_length(name, min = 0, max = 200): 42406 fails (includes all of the above)

These are probably blank strings in gene_info?

```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(	!is.na(hgnc_symbol))
cf <- confront(gene_metadata_new, v)
out <- values(cf)
is_unique_fails_hgnc <- gene_metadata_new[!out,]
is_unique_fails_hgnc
```

```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(	!is.na(name))
cf <- confront(gene_metadata_new, v)
out <- values(cf)
is_unique_fails_name <- gene_metadata_new[!out,]
is_unique_fails_name
```

```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(	!is.na(summary))
cf <- confront(gene_metadata_new, v)
out <- values(cf)
is_unique_fails_summary <- gene_metadata_new[!out,]
is_unique_fails_summary
```


```{r}
setdiff(is_unique_fails_hgnc, is_unique_fails_name)
setdiff(is_unique_fails_hgnc, is_unique_fails_summary)
```

***
# Cross-file validation with gene_info
***

Find all diffs

```{r}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_gene_metadata_info.R")

gene_info <- gene_info_new[order(gene_info_new$ensembl_gene_id),]
gene_metadata <- gene_metadata_new[order(gene_metadata_new$ensembl_gene_id),]

nrow(gene_info) == nrow(gene_metadata)

merged <- merge(x = gene_info, y = gene_metadata, by = "ensembl_gene_id", all = TRUE)

test(merged, rules_gene_metadata_info)
```


```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(name.x == name.y)
cf <- confront(merged, v)
out <- values(cf)
is_unique_fails <- merged[!out,]
is_unique_fails
```


Changed booleans that didn't propagate

```{r}
ensgs <- gene_metadata_new$ensembl_gene_id
rna_studied <- gene_metadata_new$rna_brain_change_studied
rna_changed <- gene_metadata_new$rna_in_ad_brain_change
protein_studied <- gene_metadata_new$protein_brain_change_studied
protein_changed <- gene_metadata_new$protein_in_ad_brain_change

metadata = data.frame(ensembl_gene_id = ensgs, 
                rna_studied = rna_studied, 
                rna_changed = rna_changed, 
                protein_studied = protein_studied, 
                protein_changed = protein_changed)

info_ensgs <- gene_info_new$ensembl_gene_id
info_rna_studied <- gene_info_new$rna_brain_change_studied
info_rna_changed <- gene_info_new$isAnyRNAChangedInADBrain
info_protein_studied <- gene_info_new$protein_brain_change_studied
info_protein_changed <- gene_info_new$isAnyProteinChangedInADBrain

info = data.frame(ensembl_gene_id = info_ensgs, 
                rna_studied = info_rna_studied, 
                rna_changed = info_rna_changed, 
                protein_studied = info_protein_studied, 
                protein_changed = info_protein_changed)


metadata[metadata$ensembl_gene_id == 'ENSG00000142892',]
info[info$ensembl_gene_id == 'ENSG00000142892',]

metadata[metadata$ensembl_gene_id == 'ENSG00000233009',]
info[info$ensembl_gene_id == 'ENSG00000233009',]

metadata[metadata$ensembl_gene_id == 'ENSG00000200882',]
info[info$ensembl_gene_id == 'ENSG00000200882',]
```