---
title: "Model AD Explorer"
---

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```


--------------
Pathology data
--------------

Where do the tissue values in the dropdown come from?
-> Codes appended to the specimenId
TODO: Are these consistent across datasets?

Download source files
```{r}
# GFAP, S100B
fivex_acd <- download_file('syn22049816', type = "csv")
# THIOS, Iba1
fivex_mcd <- download_file('syn22049825', type = "csv")
# LAMP1
fivex_dn <- download_file('syn22049817', type = "csv")


# GFAP, S100B, ThioS, IBA1, LAMP1, AT8, HT7 (unused: pT217, PV, WFA, polyTau, silver stain NFT)
threex_multi <- download_file('syn26601504', type = "csv")

# Soluble Abeta 40/42
fivex_sabeta <- download_file('syn22101767', type = "csv")
threex_sabeta <- download_file('syn25913411', type = "csv") 

# Insoluble Abeta 40/42
fivex_iabeta <- download_file('syn22101766', type = "csv")
threex_iabeta <- download_file('syn25913418', type = "csv") 
```

```{r}
unique(threex_multi$stain)
```

Astrocyte Cell Density GFAP
```{r}
GFAP_5x <- fivex_acd[fivex_acd$Stain == "GFAP",]
GFAP_3x <- threex_multi[threex_multi$stain == "GFAP",]
print(GFAP_5x)
print(GFAP_3x)
```

Astrocyte Cell Density S100B
```{r}
S100B_5x <- fivex_acd[fivex_acd$Stain == "S100B",]
S100B_3x <- threex_multi[threex_multi$stain == "S100B",]
print(S100B_5x)
print(S100B_3x)
```

Microglial Cell Density (IBA1)
```{r}

# thios, iba1, s100b, at8, ht7, lamp1
threex_mcd <- download_file('syn26601504', type = "csv") # file type: csv, tsv, json, f (feather)
```


```{r}
IBA1_5x <- fivex_mcd[fivex_mcd$Stain == "Iba1",]
IBA1_3x <- threex_mcd[threex_mcd$stain == "IBA1",]
print(IBA1_5x)
print(IBA1_3x)
```

Phosphotau
```{r}
AT8_3x <- threex_mcd[threex_mcd$stain == "AT8",]
print(AT8_3x)
```
Plaque Density (Thio-S)
```{r}
Thio_5x <- fivex_mcd[fivex_mcd$Stain == "ThioS",]
Thio_3x <- threex_mcd[threex_mcd$stain == "ThioS",]
print(Thio_5x)
print(Thio_3x)
```


Soluble Abetas
```{r}
fivex_sab <- download_file('syn22101767', type = "csv")
```

```{r}
individual <- fivex_sab[fivex_sab$individualID == 4255,]
print(individual)
```

Biospecimen metadaa
```{r}
pathology_biospecimen <- download_file('syn18876530.9', type = "csv")
```
```{r}
pathology_biospecimen
```
Why don't we use the tissue column from biospecimen?
```{r}
tissue_hippo <- pathology_biospecimen[pathology_biospecimen$tissue == "hippocampus",]
specimen_h <- pathology_biospecimen[str_detect(pathology_biospecimen$specimenID, "h"),]
specimen_H <- pathology_biospecimen[str_detect(pathology_biospecimen$specimenID, "H"),]
specimen_hippo <- rbind(specimen_h, specimen_H)

nrow(tissue_hippo)
nrow(specimen_hippo)
anti_join(specimen_hippo, tissue_hippo)
anti_join(tissue_hippo, specimen_hippo)

tissue_cortex <- pathology_biospecimen[pathology_biospecimen$tissue == "cerebral cortex",]
specimen_c <- pathology_biospecimen[str_detect(pathology_biospecimen$specimenID, "c"),]
specimen_C <- pathology_biospecimen[str_detect(pathology_biospecimen$specimenID, "C"),]
specimen_cortex <- rbind(specimen_c, specimen_C)

nrow(tissue_cortex)
nrow(specimen_cortex)
anti_join(specimen_cortex, tissue_cortex)
anti_join(tissue_cortex, specimen_cortex)

tissue_plasma <- pathology_biospecimen[pathology_biospecimen$tissue == "plasma",]
specimen_p <- pathology_biospecimen[str_detect(pathology_biospecimen$specimenID, "p"),]
specimen_P <- pathology_biospecimen[str_detect(pathology_biospecimen$specimenID, "P"),]
specimen_plasma <- rbind(specimen_p, specimen_P)

nrow(tissue_plasma)
nrow(specimen_plasma)
anti_join(specimen_plasma, tissue_plasma)
anti_join(tissue_plasma, specimen_plasma)
```

Individual metadata
```{r}
pathology_individual <- download_file('syn18880070.12', type = "csv")
pathology_individual
```


```{r}

```


```{r}
path_5x
```

--------------------
Gene Expression data 
--------------------


```{r}
fivex_f_12mo <- download_file('syn24847795', type = "csv") 
```


```{r}
print(fivex_f_12mo)
```

Where do gene names come from? The data files contain only ENMSUG values.
-> gene_expressions_labels.rda

What does gene_expression_labels.rda contain?

```{r}
path <- '/Users/jbritton/IdeaProjects/magora/data/gene_expressions_labels.rda'
expression_labels <- load(path)
```

```{r}
gene_expressions_labels
```
Is this the same gene expression data joined to names?
-> I think so, the l2fc value matches across one strain/sex/age for hemibrain (cortex and hippocampus are from other source files)
? is this joined to the expression data somehow? 

```{r}
strain <- gene_expressions_labels[gene_expressions_labels$mouse_model == '5xFAD', ]
strain_sex <- strain[strain$sex == 'Female',]
strain_sex_age <- strain_sex[strain_sex$age == 12, ]
print(strain_sex_age)
print(strain_sex_age[strain_sex_age$gene == 'Ptpn6',])
print(fivex_f_12mo[fivex_f_12mo$GeneID == 'ENSMUSG00000004266',])
```

Are the columns consistent across all of the gene expression files?

```{r}
path <- '/Users/jbritton/IdeaProjects/magora/data-raw/gene_expressions/data_sources.csv'
expression_data_sources <- read.csv(path)
```


```{r}
synIds <- matrix(, 0, ncol = 1)

for(i in 1:nrow(expression_data_sources)) {
  syn <- expression_data_sources[i,]$id
  version <- expression_data_sources[i,]$version
  id <- paste(syn, version, sep = '.')
  synIds <- rbind(synIds, id)
}

for(i in 1:nrow(expression_data_sources)) {
  data <- download_file(synIds[i], type = 'csv')
  assign(synIds[i], data, envir = .GlobalEnv)
}

```


```{r}
cols <- matrix(, 0, ncol = 8)

for(i in 1:nrow(expression_data_sources)) {
  col_names <- colnames(get(synIds[i]))
  if(length(col_names) != 8) {
    print("ERROR!", synId, " has the wrong number of columns!")
  }
  cols <- rbind(cols, col_names)
}
#print(cols)

print(unique(cols[,1]))
print(unique(cols[,2]))
print(unique(cols[,3]))
print(unique(cols[,4]))
print(unique(cols[,5]))
print(unique(cols[,6]))
print(unique(cols[,7]))
print(unique(cols[,8]))
```


----------
Nanostring
----------
Individual & biospecimen metadata
```{r}
nano_individual <- download_file('syn22107818.7', type = "csv") 
nano_biospecimen <- download_file('syn22107820.6', type = "csv") 

nano_individual
nano_biospecimen
```



Source file with mouse results data
```{r}
nano_mouse_source <- download_file('syn22105392.2', type = "csv") 
nano_mouse_source
```

Human results source file
```{r}
nano_human_source <- download_file('syn14237651.1', type = "tsv") 
nano_human_source
```

How do we join human and mouse data??
toupper(mouse$X) <-> human$hgnc_symbol
```{r}
nano_mouse_source_toupper <- nano_mouse_source %>% mutate(X = toupper(X))

nano_mouse_genes <- nano_mouse_source_toupper$X
nano_human_genes <- nano_human_source$hgnc_symbol


intersect(nano_mouse_genes, nano_human_genes)

```


What's in syn21483261 (aggregateModules.rda)?


```{r}
path <- '/Users/jbritton/Downloads/aggregateModules.rda'
syn_data <- load(path)
print(syn_data) # mainfest of names of all the loaded objects

# modules from nanostring.R
code_data <- c("cbe_mods", "dlpfc_mods", "fp_mods", "ifg_mods", "phg_mods", "stg_mods", "tcx_mods")
print(code_data)

setdiff(syn_data, code_data)

```

Modules

```{r}
ampad_modules_raw <- map_dfr(code_data, ~ get(.x)[["df"]]) %>%
  as_tibble()

# useful ways to peek into this thing
names(ampad_modules_raw)
names(dlpfc_mods$mods)
```

```{r}
unique(ampad_modules_raw$Module)
print(ampad_modules_raw[ampad_modules_raw$Module == "CBEturquoise",])
print(ampad_modules_raw[ampad_modules_raw$Module == "CBEblue",])
print(ampad_modules_raw[ampad_modules_raw$Module == "CBEbrown",])
print(ampad_modules_raw[ampad_modules_raw$Module == "CBEyellow",])
print(ampad_modules_raw[ampad_modules_raw$Module == "DLPFCturquoise",])
print(ampad_modules_raw[ampad_modules_raw$Module == "DLPFCblue",])
print(ampad_modules_raw[ampad_modules_raw$Module == "DLPFCbrown",])
print(ampad_modules_raw[ampad_modules_raw$Module == "DLPFCyellow",])
print(ampad_modules_raw[ampad_modules_raw$Module == "FPturquoise",])
print(ampad_modules_raw[ampad_modules_raw$Module == "FPblue",])
print(ampad_modules_raw[ampad_modules_raw$Module == "FPbrown",])
print(ampad_modules_raw[ampad_modules_raw$Module == "FPyellow",])
print(ampad_modules_raw[ampad_modules_raw$Module == "IFGturquoise",])
print(ampad_modules_raw[ampad_modules_raw$Module == "IFGblue",])
print(ampad_modules_raw[ampad_modules_raw$Module == "IFGbrown",])
print(ampad_modules_raw[ampad_modules_raw$Module == "IFGyellow",])
print(ampad_modules_raw[ampad_modules_raw$Module == "PHGturquoise",])
print(ampad_modules_raw[ampad_modules_raw$Module == "PHGblue",])
print(ampad_modules_raw[ampad_modules_raw$Module == "PHGbrown",])
print(ampad_modules_raw[ampad_modules_raw$Module == "PHGyellow",])
print(ampad_modules_raw[ampad_modules_raw$Module == "PHGgreen",])
print(ampad_modules_raw[ampad_modules_raw$Module == "STGturquoise",])
print(ampad_modules_raw[ampad_modules_raw$Module == "STGblue",])
print(ampad_modules_raw[ampad_modules_raw$Module == "STGbrown",])
print(ampad_modules_raw[ampad_modules_raw$Module == "STGyellow",])
print(ampad_modules_raw[ampad_modules_raw$Module == "TCXturquoise",])
print(ampad_modules_raw[ampad_modules_raw$Module == "TCXblue",])
print(ampad_modules_raw[ampad_modules_raw$Module == "TCXbrown",])
print(ampad_modules_raw[ampad_modules_raw$Module == "TCXyellow",])
print(ampad_modules_raw[ampad_modules_raw$Module == "TCXgreen",])
```

```{r}
ampad_modules_raw <- map_dfr("dlpfc_mods", ~ get(.x)[["df"]]) %>%
  as_tibble()

unique(ampad_modules_raw$Module)
unique(ampad_modules_raw$method)
unique(ampad_modules_raw$ModuleName)
unique(ampad_modules_raw$ModuleNameFull)
print(ampad_modules_raw)
```


```{r}
moduleGraph <- dlpfc_mods$moduleGraph
print(moduleGraph)
```


New source file validation
MG-14 - aggretageModules

old rda: syn21483261
new csv: syn51515825

```{r}
aggregate_new <- download_file('syn51515825', type = "csv") 
```


```{r}
aggregate_new
length(unique(aggregate_new$GeneID))
length(unique(aggregate_new$Module))
length(unique(aggregate_new$brainRegion))
length(unique(aggregate_new$external_gene_name))

sum(is.na(aggregate_new$GeneID))
sum(is.na(aggregate_new$external_gene_name))
```


```{r}
nrow(ampad_modules_raw[ampad_modules_raw$Module == "CBEturquoise",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "CBEblue",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "CBEbrown",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "CBEyellow",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "DLPFCturquoise",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "DLPFCblue",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "DLPFCbrown",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "DLPFCyellow",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "FPturquoise",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "FPblue",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "FPbrown",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "FPyellow",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "IFGturquoise",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "IFGblue",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "IFGbrown",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "IFGyellow",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "PHGturquoise",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "PHGblue",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "PHGbrown",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "PHGyellow",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "PHGgreen",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "STGturquoise",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "STGblue",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "STGbrown",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "STGyellow",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "TCXturquoise",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "TCXblue",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "TCXbrown",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "TCXyellow",]) +
nrow(ampad_modules_raw[ampad_modules_raw$Module == "TCXgreen",])
```

```{r}
CBEturquoise_module <- (ampad_modules_raw[ampad_modules_raw$Module == "CBEturquoise",])
CBEblue_module <- (ampad_modules_raw[ampad_modules_raw$Module == "CBEblue",])
CBEbrown_module <- (ampad_modules_raw[ampad_modules_raw$Module == "CBEbrown",])
CBEyellow_module <- (ampad_modules_raw[ampad_modules_raw$Module == "CBEyellow",])
DLPFCturquoise_module <- (ampad_modules_raw[ampad_modules_raw$Module == "DLPFCturquoise",])
DLPFCblue_module <- (ampad_modules_raw[ampad_modules_raw$Module == "DLPFCblue",])
DLPFCbrown_module <- (ampad_modules_raw[ampad_modules_raw$Module == "DLPFCbrown",])
DLPFCyellow_module <- (ampad_modules_raw[ampad_modules_raw$Module == "DLPFCyellow",])
FPturquoise_module <- (ampad_modules_raw[ampad_modules_raw$Module == "FPturquoise",])
FPblue_module <- (ampad_modules_raw[ampad_modules_raw$Module == "FPblue",])
FPbrown_module <- (ampad_modules_raw[ampad_modules_raw$Module == "FPbrown",])
FPyellow_module <- (ampad_modules_raw[ampad_modules_raw$Module == "FPyellow",])
IFGturquoise_module <- (ampad_modules_raw[ampad_modules_raw$Module == "IFGturquoise",])
IFGblue_module <- (ampad_modules_raw[ampad_modules_raw$Module == "IFGblue",])
IFGbrown_module <- (ampad_modules_raw[ampad_modules_raw$Module == "IFGbrown",])
IFGyellow_module <- (ampad_modules_raw[ampad_modules_raw$Module == "IFGyellow",])
PHGturquoise_module <- (ampad_modules_raw[ampad_modules_raw$Module == "PHGturquoise",])
PHGblue_module <- (ampad_modules_raw[ampad_modules_raw$Module == "PHGblue",])
PHGbrown_module <- (ampad_modules_raw[ampad_modules_raw$Module == "PHGbrown",])
PHGyellow_module <- (ampad_modules_raw[ampad_modules_raw$Module == "PHGyellow",])
PHGgreen_module <- (ampad_modules_raw[ampad_modules_raw$Module == "PHGgreen",])
STGturquoise_module <- (ampad_modules_raw[ampad_modules_raw$Module == "STGturquoise",])
STGblue_module <- (ampad_modules_raw[ampad_modules_raw$Module == "STGblue",])
STGbrown_module <- (ampad_modules_raw[ampad_modules_raw$Module == "STGbrown",])
STGyellow_module <- (ampad_modules_raw[ampad_modules_raw$Module == "STGyellow",])
TCXturquoise_module <- (ampad_modules_raw[ampad_modules_raw$Module == "TCXturquoise",])
TCXblue_module <- (ampad_modules_raw[ampad_modules_raw$Module == "TCXblue",])
TCXbrown_module <- (ampad_modules_raw[ampad_modules_raw$Module == "TCXbrown",])
TCXyellow_module <- (ampad_modules_raw[ampad_modules_raw$Module == "TCXyellow",])
TCXgreen_module <- (ampad_modules_raw[ampad_modules_raw$Module == "TCXgreen",])
```



