---
title: "misc investigations etc."
---
*****
Setup
*****
```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```


*****
MSSM/ZHANG
*****
additional notes on https://sagebionetworks.jira.com/browse/AG-1252

prior to loading csv into this workbook:
1. populate internal-only columns (Source, input_data, initial_nomination_)
2. adjust user-supplied values as required (study, Team)
3. export each sheet as csv
4. remove 'empty' rows from csvs as required (new only for this one)

In this workbook:
1. remove genes in 'updates' from 'new' 
2. merge Predicted_Theraputic_Direction_* columns -> Predicted_Theraputic_Direction
3. populate 'internal' columns (source, intital_nomination)
4. export modified csv that can be cleanly merged with targets source file

```{r}
zhang_new <- read_csv("/Users/jbritton/Desktop/New nominations/from drive/Zhang/Zhang_new_noms.csv")

zhang_updates <- read_csv("/Users/jbritton/Desktop/New nominations/from drive/Zhang/Zhang_updated_noms.csv")

zhang_new
zhang_updates
```
Sanity check uniques
new - no dups
updates - 3 dups?? -> TODO why????!?!?!?!?!
```{r}
# are there dups?
length(unique(zhang_new$ensembl_gene_id))
nrow(zhang_new)

length(unique(zhang_updates$ensembl_gene_id))
nrow(zhang_updates)

# what are the dups?
n_occur <- data.frame(table(zhang_updates$ensembl_gene_id))
dup_updates <- zhang_updates[zhang_updates$ensembl_gene_id %in% n_occur$Var1[n_occur$Freq > 1],]

# are the dup rows identical? -> Nope
highlander <- function(x) {
  x |> 
    unique() |> 
    length() |> 
    {\(x) x == 1}()
}

dup_updates |> 
  group_by(ensembl_gene_id) |> 
  summarise(
    across(everything(), highlander)
  )

# look at the dups
zhang_updates[zhang_updates$ensembl_gene_id == "ENSG00000118733",]
zhang_updates[zhang_updates$ensembl_gene_id == "ENSG00000124785",]
zhang_updates[zhang_updates$ensembl_gene_id == "ENSG00000147065",]
```

remove targets on the updates list from the new list, because they are not in fact new
```{r}
keep <- zhang_new[!(zhang_new$ensembl_gene_id %in% zhang_updates$ensembl_gene_id),]

dropped <- zhang_new[zhang_new$ensembl_gene_id %in% zhang_updates$ensembl_gene_id,]
cat(dropped$hgnc_symbol)
cat("\nnew: ", nrow(zhang_new), "\nupdates: ", nrow(zhang_updates), "\ndropped: ", nrow(dropped), "\nkeeps: ", nrow(keep))
keep
dropped
```


Merge Predicted_Therapeutic_Direction_* fields in 'keep'
```{r}
merged_ptd <- paste(keep$`Predicted_Therapeutic_Direction category`, keep$`Predicted_Therapeutic_Direction details`)
merged_ptd_keeps <- keep$`Predicted_Therapeutic_Direction category` <- merged_ptd
drop_cols <- c("Predicted_Therapeutic_Direction details")
merged_ptd_keeps <- keep[, !(names(keep) %in% drop_cols)]
names(merged_ptd_keeps)[names(merged_ptd_keeps) == "Predicted_Therapeutic_Direction category"] <- "Predicted_Therapeutic_Direction"
merged_ptd_keeps
```

Write new nominations csv file
```{r}
write.csv(merged_ptd_keeps, "/Users/jbritton/Desktop/New nominations/from drive/Zhang/Zhang_new_noms_deduped.csv", row.names=FALSE)

```


*****
Emory
*****
additional notes on https://sagebionetworks.jira.com/browse/AG-1254

prior to loading csv into this workbook:
1. populate internal-only columns (Source, input_data, initial_nomination)
2. adjust user-supplied values as required (Study, Team)
3. export each sheet as csv
4. remove 'empty' rows from csvs as required

In this workbook:
1. merge Predicted_Theraputic_Direction_* columns -> Predicted_Theraputic_Direction
2. export modified csv that can be cleanly merged with targets source file

```{r}
emory_new <- read_csv("/Users/jbritton/Desktop/New nominations/from drive/Emory/12.6.23_Emory Agora Target Submission filled+11Nov23_updates.csv")

emory_updates <- read_csv("/Users/jbritton/Desktop/New nominations/from drive/Emory/Emory Agora target submissions_for_empty_updates.xlsm - Nominated Target Updates.csv")

emory_new
emory_updates
```

Sanity check uniques
new - 
updates - 
```{r}
# are there dups?
length(unique(emory_new$ensembl_gene_id))
nrow(emory_new)

length(unique(emory_updates$ensembl_gene_id))
nrow(emory_updates)

# what are the dups?
n_occur <- data.frame(table(emory_updates$ensembl_gene_id))
dup_updates <- emory_updates[emory_updates$ensembl_gene_id %in% n_occur$Var1[n_occur$Freq > 1],]

# are the dup rows identical? -> Nope
highlander <- function(x) {
  x |> 
    unique() |> 
    length() |> 
    {\(x) x == 1}()
}

dup_updates |> 
  group_by(ensembl_gene_id) |> 
  summarise(
    across(everything(), highlander)
  )



```



Merge Predicted_Therapeutic_Direction_* fields
```{r}
merged_ptd <- paste(emory_new$`Predicted_Therapeutic_Direction category`, emory_new$`Predicted_Therapeutic_Direction details`)
merged_ptd_keeps <- emory_new$`Predicted_Therapeutic_Direction category` <- merged_ptd
drop_cols <- c("Predicted_Therapeutic_Direction details")
merged_ptd_keeps <- emory_new[, !(names(emory_new) %in% drop_cols)]
names(merged_ptd_keeps)[names(merged_ptd_keeps) == "Predicted_Therapeutic_Direction category"] <- "Predicted_Therapeutic_Direction"
merged_ptd_keeps
```

Write new nominations csv file
```{r}
write.csv(merged_ptd_keeps, "/Users/jbritton/Desktop/New nominations/from drive/Emory/Emory_ptd_merged.csv", row.names=FALSE)
```