---
title: "distribution_data.json"
---

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```

# Validation run metadata
```{r echo=FALSE}
# add file-specific values here
name <- "distribution_data"
source_synId <- "syn25575156.13" # new 5 point overall score introduced
old_synId <-  "syn26274945" 
new_synId <- "syn27572407"

# score-type specific rules files
genetics_rules <- paste0("rules_", name, "_genetics")
omics_rules <- paste0("rules_", name, "_omics")
lit_rules <- paste0("rules_", name, "_lit")
overall_rules <- paste0("rules_", name, "_overall")

# not file-specific
old_name <- paste(name, "old", sep="_")
new_name <- paste(name, "new", sep="_")
rules <- paste0("rules_", name) # unused for this file

# print validation summary
cat( name, ".json", "\n", date(),
    "\nold: ", old_synId, " \nnew: ", new_synId,
    sep="")
```

# Download files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
download_files(old_synId, new_synId, name)
```
```{r}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source <- download_file(source_synId, type = "table")
assign(paste(name, "source", sep="_"), source, envir = .GlobalEnv)
```

# Compared old & new files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
compare(get(old_name), get(new_name), name)
```


# Validated new file against rules
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_distribution_data_genetics.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_distribution_data_omics.R")
# source("~/repos/rstudio/agora-data-validation/src/rules/rules_distribution_data_lit.R") Removed this score in 3.2.0 (post m49 data release)
source("~/repos/rstudio/agora-data-validation/src/rules/rules_distribution_data_overall.R")

# extract nested score-specific distribution data as dataframes
GeneticsScores <- distribution_data_new[, "genetics_score"]
OmicsScores <- distribution_data_new[, "multi_omics_score"]
# LitScores <- distribution_data_new[, "literature_score"]
OverallScores <- distribution_data_new[, "target_risk_score"]

test(GeneticsScores, get(genetics_rules))
test(OmicsScores, get(omics_rules))
# test(LitScores, get(lit_rules)) 
test(OverallScores, get(overall_rules))
```

***
# Outstanding validation failures
***
N/A

***
# Validation failure investigations
***

```{r echo=FALSE}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(VALIDATION_RULE)
cf <- confront(DATAFRAME, v)
out <- values(cf)
is_unique_fails <- DATAFRAME[!out,]
is_unique_fails
```


***
# Other validation
***

## Lit score empty top bin issue (See AG-827 comments)
The transform logic was returning distributions sorted by value (highest to lowest)
The application was assuming the distributions were sorted by bin ordering
I think that this means we have been presenting inaccurate histograms in Agora for all score types :(
The solution is to add a sort=false parameter to the value_counts call in calculate_distribution

=> The inappropriate sort was removed via AG-827

# Generate distributions locally from source data and check against processed results

Expected distributions:
```{r}
overall_only <- distribution_data_source %>% select(ENSG, Overall)
overall_only[nrow(overall_only) + 1,] = c("max", 5)
overall_only$Overall <- as.numeric(overall_only$Overall)

genetics_scored <- distribution_data_source[distribution_data_source$isScored_genetics == 'Y',]
genetics_only <- genetics_scored %>% select(ENSG, GeneticsScore)
genetics_only[nrow(genetics_only) + 1,] = c("max", 3)
genetics_only$GeneticsScore <- as.numeric(genetics_only$GeneticsScore)

omics_scored <- distribution_data_source[distribution_data_source$isScored_omics == 'Y',]
omics_only <- omics_scored %>% select(ENSG, OmicsScore)
omics_only[nrow(omics_only) + 1,] = c("max", 2)
omics_only$OmicsScore <- as.numeric(omics_only$OmicsScore)

lit_scored <- distribution_data_source[distribution_data_source$isScored_lit == 'Y',]
lit_only <- lit_scored %>% select(ENSG, LiteratureScore)
lit_only[nrow(lit_only) + 1,] = c("max", 2)
lit_only$LiteratureScore <- as.numeric(lit_only$LiteratureScore)

overall_binned <- overall_only %>% mutate(overall_bin = cut(Overall, breaks=10))
genetics_binned <- genetics_only %>% mutate(genetics_bin = cut(GeneticsScore, breaks=10))
omics_binned <- omics_only %>% mutate(omics_bin = cut(OmicsScore, breaks=10))
lit_binned <- lit_only %>% mutate(lit_bin = cut(LiteratureScore, breaks=10))

# expected no genes per bin
overall_bins <- unique(overall_binned$overall_bin)
cat('\nOverall\n')
cat('bin 1: ', sum(overall_binned$overall_bin == '(-0.005,0.5]'), '\n')
cat('bin 2: ', sum(overall_binned$overall_bin == '(0.5,1]'), '\n')
cat('bin 3: ', sum(overall_binned$overall_bin == '(1,1.5]'), '\n')
cat('bin 4: ', sum(overall_binned$overall_bin == '(1.5,2]'), '\n')
cat('bin 5: ', sum(overall_binned$overall_bin == '(2,2.5]'), '\n')
cat('bin 6: ', sum(overall_binned$overall_bin == '(2.5,3]'), '\n')
cat('bin 7: ', sum(overall_binned$overall_bin == '(3,3.5]'), '\n')
cat('bin 8: ', sum(overall_binned$overall_bin == '(3.5,4]'), '\n')
cat('bin 9: ', sum(overall_binned$overall_bin == '(4,4.5]'), '\n')
cat('bin 10: ', sum(overall_binned$overall_bin == '(4.5,5]')- 1, '\n')

genetics_bins <- unique(genetics_binned$genetics_bin)
cat('\nGenetics\n')
cat('bin 1: ', sum(genetics_binned$genetics_bin == '(-0.003,0.3]'), '\n')
cat('bin 2: ', sum(genetics_binned$genetics_bin == '(0.3,0.6]'), '\n')
cat('bin 3: ', sum(genetics_binned$genetics_bin == '(0.6,0.9]'), '\n')
cat('bin 4: ', sum(genetics_binned$genetics_bin == '(0.9,1.2]'), '\n')
cat('bin 5: ', sum(genetics_binned$genetics_bin == '(1.2,1.5]'), '\n')
cat('bin 6: ', sum(genetics_binned$genetics_bin == '(1.5,1.8]'), '\n')
cat('bin 7: ', sum(genetics_binned$genetics_bin == '(1.8,2.1]'), '\n')
cat('bin 8: ', sum(genetics_binned$genetics_bin == '(2.1,2.4]'), '\n')
cat('bin 9: ', sum(genetics_binned$genetics_bin == '(2.4,2.7]'), '\n')
cat('bin 10: ', sum(genetics_binned$genetics_bin == '(2.7,3]')- 1, '\n')

omics_bins <- unique(omics_binned$omics_bin)
cat('\nOmics\n')
cat('bin 1: ', sum(omics_binned$omics_bin == '(-0.002,0.2]'), '\n')
cat('bin 2: ', sum(omics_binned$omics_bin == '(0.2,0.4]'), '\n')
cat('bin 3: ', sum(omics_binned$omics_bin == '(0.4,0.6]'), '\n')
cat('bin 4: ', sum(omics_binned$omics_bin == '(0.6,0.8]'), '\n')
cat('bin 5: ', sum(omics_binned$omics_bin == '(0.8,1]'), '\n')
cat('bin 6: ', sum(omics_binned$omics_bin == '(1,1.2]'), '\n')
cat('bin 7: ', sum(omics_binned$omics_bin == '(1.2,1.4]'), '\n')
cat('bin 8: ', sum(omics_binned$omics_bin == '(1.4,1.6]'), '\n')
cat('bin 9: ', sum(omics_binned$omics_bin == '(1.6,1.8]'), '\n')
cat('bin 10: ', sum(omics_binned$omics_bin == '(1.8,2]')- 1,  '\n')


lit_bins <- unique(lit_binned$lit_bin)
cat('\nLiterature\n')
cat('bin 1: ', sum(lit_binned$lit_bin == '(-0.002,0.2]'), '\n')
cat('bin 2: ', sum(lit_binned$lit_bin == '(0.2,0.4]'), '\n')
cat('bin 3: ', sum(lit_binned$lit_bin == '(0.4,0.6]'), '\n')
cat('bin 4: ', sum(lit_binned$lit_bin == '(0.6,0.8]'), '\n')
cat('bin 5: ', sum(lit_binned$lit_bin == '(0.8,1]'), '\n')
cat('bin 6: ', sum(lit_binned$lit_bin == '(1,1.2]'), '\n')
cat('bin 7: ', sum(lit_binned$lit_bin == '(1.2,1.4]'), '\n')
cat('bin 8: ', sum(lit_binned$lit_bin == '(1.4,1.6]'), '\n')
cat('bin 9: ', sum(lit_binned$lit_bin == '(1.6,1.8]'), '\n')
cat('bin 10: ', sum(lit_binned$lit_bin == '(1.8,2]')- 1, '\n')
```

Did this impact our old distributions too?
```{r}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source <- download_file("syn25575156.9", type = "table")
assign(paste(name, "source_old", sep="_"), source, envir = .GlobalEnv)
```

Expected old distributions:
The expected distributions do not match what is in the app today, so same bug :(
```{r}
overall_fn <- distribution_data_source_old %>% select(ENSG, Overall, FlyNeuroPathScore)
overall_fn$Overall <- overall_fn$Overall - overall_fn$FlyNeuroPathScore
overall_only <- overall_fn %>% select(ENSG, Overall)
overall_only[nrow(overall_only) + 1,] = c("max", 7)
overall_only[nrow(overall_only) + 1,] = c("min", 0)
overall_only$Overall <- as.numeric(overall_only$Overall)

genetics_scored <- distribution_data_source_old[distribution_data_source_old$isScored_genetics == 'Y',]
genetics_only <- genetics_scored %>% select(ENSG, GeneticsScore)
genetics_only[nrow(genetics_only) + 1,] = c("max", 3)
genetics_only[nrow(genetics_only) + 1,] = c("min", 0)
genetics_only$GeneticsScore <- as.numeric(genetics_only$GeneticsScore)

omics_scored <- distribution_data_source_old[distribution_data_source_old$isScored_omics == 'Y',]
omics_only <- omics_scored %>% select(ENSG, OmicsScore)
omics_only[nrow(omics_only) + 1,] = c("max", 2)
omics_only[nrow(omics_only) + 1,] = c("min", 0)
omics_only$OmicsScore <- as.numeric(omics_only$OmicsScore)

lit_scored <- distribution_data_source_old[distribution_data_source_old$isScored_lit == 'Y',]
lit_only <- lit_scored %>% select(ENSG, LiteratureScore)
lit_only[nrow(lit_only) + 1,] = c("max", 2)
lit_only[nrow(lit_only) + 1,] = c("min", 0)
lit_only$LiteratureScore <- as.numeric(lit_only$LiteratureScore)

overall_binned <- overall_only %>% mutate(overall_bin = cut(Overall, breaks=10))
genetics_binned <- genetics_only %>% mutate(genetics_bin = cut(GeneticsScore, breaks=10))
omics_binned <- omics_only %>% mutate(omics_bin = cut(OmicsScore, breaks=10))
lit_binned <- lit_only %>% mutate(lit_bin = cut(LiteratureScore, breaks=10))

# expected no genes per bin
overall_bins <- unique(overall_binned$overall_bin)
cat('\nOverall\n')
cat('bin 1: ', sum(overall_binned$overall_bin == '(-0.007,0.7]') - 1, '\n')
cat('bin 2: ', sum(overall_binned$overall_bin == '(0.7,1.4]'), '\n')
cat('bin 3: ', sum(overall_binned$overall_bin == '(1.4,2.1]'), '\n')
cat('bin 4: ', sum(overall_binned$overall_bin == '(2.1,2.8]'), '\n')
cat('bin 5: ', sum(overall_binned$overall_bin == '(2.8,3.5]'), '\n')
cat('bin 6: ', sum(overall_binned$overall_bin == '(3.5,4.2]'), '\n')
cat('bin 7: ', sum(overall_binned$overall_bin == '(4.2,4.9]'), '\n')
cat('bin 8: ', sum(overall_binned$overall_bin == '(4.9,5.6]'), '\n')
cat('bin 9: ', sum(overall_binned$overall_bin == '(5.6,6.3]'), '\n')
cat('bin 10: ', sum(overall_binned$overall_bin == '(6.3,7.01]')- 1, '\n')

genetics_bins <- unique(genetics_binned$genetics_bin)
cat('\nGenetics\n')
cat('bin 1: ', sum(genetics_binned$genetics_bin == '(-0.003,0.3]') - 1, '\n')
cat('bin 2: ', sum(genetics_binned$genetics_bin == '(0.3,0.6]'), '\n')
cat('bin 3: ', sum(genetics_binned$genetics_bin == '(0.6,0.9]'), '\n')
cat('bin 4: ', sum(genetics_binned$genetics_bin == '(0.9,1.2]'), '\n')
cat('bin 5: ', sum(genetics_binned$genetics_bin == '(1.2,1.5]'), '\n')
cat('bin 6: ', sum(genetics_binned$genetics_bin == '(1.5,1.8]'), '\n')
cat('bin 7: ', sum(genetics_binned$genetics_bin == '(1.8,2.1]'), '\n')
cat('bin 8: ', sum(genetics_binned$genetics_bin == '(2.1,2.4]'), '\n')
cat('bin 9: ', sum(genetics_binned$genetics_bin == '(2.4,2.7]'), '\n')
cat('bin 10: ', sum(genetics_binned$genetics_bin == '(2.7,3]')- 1, '\n')

omics_bins <- unique(omics_binned$omics_bin)
cat('\nOmics\n')
cat('bin 1: ', sum(omics_binned$omics_bin == '(-0.002,0.2]') - 1, '\n')
cat('bin 2: ', sum(omics_binned$omics_bin == '(0.2,0.4]'), '\n')
cat('bin 3: ', sum(omics_binned$omics_bin == '(0.4,0.6]'), '\n')
cat('bin 4: ', sum(omics_binned$omics_bin == '(0.6,0.8]'), '\n')
cat('bin 5: ', sum(omics_binned$omics_bin == '(0.8,1]'), '\n')
cat('bin 6: ', sum(omics_binned$omics_bin == '(1,1.2]'), '\n')
cat('bin 7: ', sum(omics_binned$omics_bin == '(1.2,1.4]'), '\n')
cat('bin 8: ', sum(omics_binned$omics_bin == '(1.4,1.6]'), '\n')
cat('bin 9: ', sum(omics_binned$omics_bin == '(1.6,1.8]'), '\n')
cat('bin 10: ', sum(omics_binned$omics_bin == '(1.8,2]')- 1,  '\n')


lit_bins <- unique(lit_binned$lit_bin)
cat('\nLiterature\n')
cat('bin 1: ', sum(lit_binned$lit_bin == '(-0.002,0.2]') - 1, '\n')
cat('bin 2: ', sum(lit_binned$lit_bin == '(0.2,0.4]'), '\n')
cat('bin 3: ', sum(lit_binned$lit_bin == '(0.4,0.6]'), '\n')
cat('bin 4: ', sum(lit_binned$lit_bin == '(0.6,0.8]'), '\n')
cat('bin 5: ', sum(lit_binned$lit_bin == '(0.8,1]'), '\n')
cat('bin 6: ', sum(lit_binned$lit_bin == '(1,1.2]'), '\n')
cat('bin 7: ', sum(lit_binned$lit_bin == '(1.2,1.4]'), '\n')
cat('bin 8: ', sum(lit_binned$lit_bin == '(1.4,1.6]'), '\n')
cat('bin 9: ', sum(lit_binned$lit_bin == '(1.6,1.8]'), '\n')
cat('bin 10: ', sum(lit_binned$lit_bin == '(1.8,2]')- 1, '\n')
```



## Distribution validation

Ensure that no scores were dropped by validating that the distributions contain the expected total number of genes
note that we have far fewer lit scores with v13 source (8614 vs )

```{r echo=FALSE}
risk <- distribution_data_new$target_risk_score
cat("\nOverall distribution total genes: ", sapply(risk$distribution, sum))
cat("\nOverall total genes expected: ", "24822") #  isScored == true from overall_scores_test.json 

omics <- distribution_data_new$multi_omics_score
cat("\n\nGenomics distribution total genes: ", sapply(omics$distribution, sum))
cat("\nGenomics total genes expected: ", "23907") #  isScored == true from overall_scores_test.json

lit <- distribution_data_new$literature_score
cat("\n\nLiterature distribution total genes: ", sapply(lit$distribution, sum))
cat("\nLiterature total genes expected: ", "8614") #  isScored == true from overall_scores_test.json

genetics <- distribution_data_new$genetics
cat("\n\nGenetics distribution total genes: ", sapply(genetics$distribution, sum))
cat("\nGenetics total genes expected: ", "24314") #  isScored == true from overall_scores_test.json
```


## Hackery to work around structure change (to compare v5 to later versions)
download the 'new' file and edit it to make it a list of objects instead of one big object to allow comparison with previous versions and validate with existing rules (until we move past this awkward phase...)
```{r}
modified_path <- "/Users/jbritton/repos/rstudio/agora-data-validation/src/validate/files/distribution_data(22).json"
distribution_data_new <- as.data.frame(fromJSON(modified_path))
```


diff -> all rows are different

```{r}

sqldf('SELECT * FROM distribution_data_new EXCEPT SELECT * FROM distribution_data_old')
sqldf('SELECT * FROM distribution_data_new INTERSECT SELECT * FROM distribution_data_old')

```