---
title: "gene_info_nominatedtarget_test"
---

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```

# Validation run metadata
```{r echo=FALSE}
# file-specific
name <- "gene_info"
sub_name <- "nominatedtarget"
source_synId <- "syn12540368.36" # harmonized_targets_20220707.csv 
old_synId <-  "syn12548902.39" # m49 Live
new_synId <- "syn17015359" 

# not file-specific
old_name <- paste(name, "old", sep="_")
new_name <- paste(name, "new", sep="_")
rules <- paste0("rules_gene_info_", sub_name)

# print validation summary
cat( name, ".json", "\nReported on:  ", date(),
    "\nold: ", old_synId, " \nnew: ", new_synId,
    sep="")
```

# Downloaded files to compare
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
download_files(old_synId, new_synId, name)
```

# Download source files
```{r}
source <- download_file(source_synId, type = "csv") 
assign(paste(name, "source", sep="_"), source, envir = .GlobalEnv)
```

# Compare old & new subobjects
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
compare_subobjects(get(old_name), get(new_name), "ensembl_gene_id", "nominatedtarget")
```

# Validate subobjects against rules

```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_gene_info_nominatedtarget.R")
test_subobject(get(new_name), "nominatedtarget", get(rules))
```

***
# Outstanding validation failures
***
* !is.na(data_synapseid)/field_length(data_synapseid, min = 3, max = 25) 
49 fails -> NA values are properly handled by the app

* !is.na(data_used_to_support_target_selection)/field_length(data_used_to_support_target_selection, min = 15, max = 2000)
2 fails -> NA values TODO AG-365

* !is.na(study)/field_length(study, min = 2, max = 200)
7 fails -> NA values - see https://sagebionetworks.jira.com/browse/AG-846

* !is.na(validation_study_details)/field_length(validation_study_details, min = 0, max = 1000)
130 fails -> NA values - see https://sagebionetworks.jira.com/browse/AG-846

***
# Validation failure investigations
```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(VALIDATION_RULE)
cf <- confront(DATAFRAME, v)
out <- values(cf)
is_unique_fails <- DATAFRAME[!out,]
# is_unique_fails <- DATAFRAME[is.na(out),] # use this for nNA failures
is_unique_fails
```

## AG-533 Missing/optional NT fields 
### data_synapseid
Addressed via https://sagebionetworks.jira.com/browse/AG-567
```{r}
# yields a dataframe containing the failure cases for a specific rule
nominated_targets <- get_subobject(get(new_name), "nominatedtarget")
v <- validator(field_length(data_synapseid, min = 3, max = 25))
cf <- confront(nominated_targets, v)
out <- values(cf)
is_unique_fails <- nominated_targets[is.na(out),]
is_unique_fails
```
### data_used_to_support_target_selection
Needs to be addressed via https://sagebionetworks.jira.com/browse/AG-365
```{r}
# yields a dataframe containing the failure cases for a specific rule
nominated_targets <- get_subobject(get(new_name), "nominatedtarget")
v <- validator(field_length(data_used_to_support_target_selection, min = 15, max = 2000))
cf <- confront(nominated_targets, v)
out <- values(cf)
is_unique_fails <- nominated_targets[is.na(out),]
is_unique_fails
```
### study (nNA)
Agora omits the 'Learn more about the target nomination process' for these
```{r}
# yields a dataframe containing the failure cases for a specific rule
nominated_targets <- get_subobject(get(new_name), "nominatedtarget")
v <- validator(!is.na(study))
cf <- confront(nominated_targets, v)
out <- values(cf)
is_unique_fails <- nominated_targets[!out,]
is_unique_fails
```

### validation_study_details
Is this supposed to be displayed on the Nominations tab? It isn't for any gene currently, just in the NT table. https://sagebionetworks.jira.com/browse/AG-846
```{r}
# yields a dataframe containing the failure cases for a specific rule
nominated_targets <- get_subobject(get(new_name), "nominatedtarget")
v <- validator(	field_length(validation_study_details, min = 0, max = 1000))
cf <- confront(nominated_targets, v)
out <- values(cf)
is_unique_fails <- nominated_targets[is.na(out),]
is_unique_fails
```
***
# Misc investigations


## New TREAT-AD nominations - AG-667
```{r}
nominated_targets_new <- get_subobject(gene_info_new, "nominatedtarget")
nominated_targets_old <- get_subobject(gene_info_old, "nominatedtarget")

ensg_new <- nominated_targets_new$ensembl_gene_id
ensg_old <- nominated_targets_old$ensembl_gene_id

new_ensgs <- setdiff(ensg_new, ensg_old)
missing_ensgs <- setdiff(ensg_old, ensg_new)

new_ensgs
missing_ensgs
```

## AG-771: TREAT-AD nominations - unicode replacement character
```{r}
#ensg <- "ENSG00000127666"
#ensg <- "ENSG00000115267"
#ensg <- "ENSG00000120915"
#ensg <- "ENSG00000108771"
#ensg <- "ENSG00000079785"
ensg <- "ENSG00000116584"
new <- gene_info_new[gene_info_new$ensembl_gene_id == ensg,]
old <- gene_info_old[gene_info_old$ensembl_gene_id == ensg,]
identical(new, old)
nom_new <- new$nominatedtarget
nom_old <- old$nominatedtarget
print(nom_new)
print(nom_old)
```


## AG-674: unicode replacement character
```{r}
#ensg <- "ENSG00000089169"
#ensg <- "ENSG00000131095"
#ensg <- "ENSG00000147065"
#ensg <- "ENSG00000163884"
#ensg <- "ENSG00000025434"
ensg <- "ENSG00000131408"
new <- gene_info_new[gene_info_new$ensembl_gene_id == ensg,]
old <- gene_info_old[gene_info_old$ensembl_gene_id == ensg,]
identical(new, old)
nom_new <- new$nominatedtarget
nom_old <- old$nominatedtarget
print(nom_new)
print(nom_old)
```

# backslash testing - TODO figure out how to make this work
to hack source file to add a \ for testing:
nominated_targets$predicted_therapeutic_direction[[1]] =  "Antagonism predicted to reduce disease \u524progression. "
```{r}
# yields a dataframe containing the failure cases for a specific rule
# v <- validator(!field_format(predicted_therapeutic_direction, "*\\u0524*", type = "regex"))
# v <- validator(!field_format(predicted_therapeutic_direction, "*\\*", type = "regex"))
# v <- validator(!field_format(predicted_therapeutic_direction, "*\\\\*", type = "regex"))
# v <- validator(!field_format(predicted_therapeutic_direction, "*\\\\\\*", type = "regex"))
v <- validator(!field_format(predicted_therapeutic_direction, "*[\\]*"))

cf <- confront(nominated_targets, v)
out <- values(cf)
is_unique_fails <- nominated_targets[!out,]
# is_unique_fails <- nominated_targets[is.na(out),] # use this for nNA failures
is_unique_fails
```