---
title: "biodomains.json"
---

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```

# Validation run metadata
```{r echo=FALSE}
# add file-specific values here
name <- "genes_biodomains"
source_synId <- "syn44151254.1"
old_synId <- "syn51062522"
new_synId <- "syn51062085"

# not file-specific
source_name <- paste(name, "source", sep="_")
old_name <- paste(name, "old", sep="_")
new_name <- paste(name, "new", sep="_")
rules <- paste0("rules_", name)

# print validation metadata summary
cat( name, ".json", "\n", date(),
    "\nold: ", old_synId, " \nnew: ", new_synId,
    sep="")
```

# Download files

## download new and old json files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
download_files(old_synId, new_synId, name)
```

## download source file
```{r}
source <- download_file(source_synId, type = "csv") # file type: csv, tsv, json, f (feather)
assign(paste(name, "source", sep="_"), source, envir = .GlobalEnv)
```

# Compared old & new files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
compare(get(old_name), get(new_name), name)
compare_subobjects(get(old_name), get(new_name), "ensembl_gene_id", "gene_biodomains")
```


# Validated new file against rules
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_genes_biodomains.R")
test(get(new_name), get(rules))
test_subobject(get(new_name), "gene_biodomains", get('rules_gene_biodomains'))

# extract a flattened list of unique GO terms for go_terms list item validation
gene_biodomains_to_test <- get_subobject(get(new_name), "gene_biodomains")
go_terms_to_test <- data.frame(unique(flatten_chr(gene_biodomains_to_test$go_terms)))
names(go_terms_to_test)[1] <- "go_terms" # add colname back
test(get("go_terms_to_test"), get('rules_gene_biodomain_go_terms')) 
```


***
# Validation failure investigations
***


```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(length(go_terms) == n_gene_biodomain_terms)
cf <- confront(genes_biodomains_new, v)
out <- values(cf)
is_unique_fails <- genes_biodomains_new[!out,]
is_unique_fails
```



```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(enrichment_pct < 1)
cf <- confront(genes_biodomains_new, v)
out <- values(cf)
is_unique_fails <- genes_biodomains_new[!out,]
is_unique_fails
```

# lengths(go_terms) == n_gene_biodomain_terms - 12 fails (ATD v35)
```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(lengths(go_terms) == n_gene_biodomain_terms)
cf <- confront(genes_biodomains_new, v)
out <- values(cf)
is_unique_fails <- genes_biodomains_new[!out,]
is_unique_fails
```
```{r}
gb <- genes_biodomains_new$gene_biodomains
results <- c()

for (g in gb) {
  r <- as.data.frame(g)
  print(r)
  x <- lengths(r$go_terms)
  print(x)
  y <- r$n_gene_biodomain_terms
  results <- x == y
}

results
```


```{r}
r <- as.data.frame(gb[1])
x <- lengths(gb1$go_terms)
y <- gb1$n_gene_biodomain_terms

x == y
```


***
# Other investigations
***
max # go terms in genes_biodomains?
```{r}
gene_biodomains <- unnest(genes_biodomains_new, cols=c(gene_biodomains))
gos <- gene_biodomains$go_terms

vals <- c(0)

for(i in 1:length(gos)) {
  n_gos <- length(gos[[i]])
  append(vals, n_gos)
  if (n_gos > 40) {
      cat("\n", n_gos)
  }

}
```


max # go terms per biodomain?
```{r}
biodomains <- unique(genes_biodomains_source$Biodomain)

for(biodomain in biodomains) {
  terms <- genes_biodomains_source[genes_biodomains_source$Biodomain == biodomain,]
  cat("\n", length(unique(terms$GOterm_Name)))
}

```

# 3.3.0 ALDmv61 validation
```{r}
genes_biodomains_new[1,]$gene_biodomains
genes_biodomains_old[1,]$gene_biodomains

```

How many unmapped genes do we have in Agora?
```{r}
gb_ensgs <- genes_biodomains_new$ensembl_gene_id
gi_ensgs <- gene_info_new$ensembl_gene_id

extras <- setdiff(gb_ensgs, gi_ensgs)
missing <- setdiff(gi_ensgs, gb_ensgs)
print("Total ENSGS (bd)")
print(length(gb_ensgs))
print("Total ENSGS (gi)")
print(length(gi_ensgs))
print("Extras (in bd not gi)")
print(length(extras))
#print(extras)
print("Missing (in gi not bd)")
print(length(missing))
#print(missing)

gi_not_bd_missings <- gene_info_new[gene_info_new$ensembl_gene_id %in% missing, ]
gi_not_bd_no_hgnc <- gi_not_bd_missings[is.na(gi_not_bd_missings$hgnc_symbol),]

print("gi NA hgnc (missing)")
print(nrow(gi_not_bd_no_hgnc))

gi_no_hgnc <- gene_info_new[is.na(gene_info_new$hgnc_symbol),]

print("gi NA hgnc (all)")
print(nrow(gi_no_hgnc))
```

ENSG matches between scores and biodomains
```{r}
scores_ensg <- overall_scores_new$ensembl_gene_id

extras <- setdiff(gb_ensgs, scores_ensg)
missing <- setdiff(scores_ensg, gb_ensgs)
print("Total ENSGS (bd)")
print(length(gb_ensgs))
print("Total ENSGS (os)")
print(length(scores_ensg))
print("Extras (in bd not os)")
print(length(extras))
#print(extras)
print("Missing (in os not bd)")
print(length(missing))
#print(missing)
print("intersect")
intersect_ensgs <- intersect(scores_ensg, gb_ensgs)
print(length(intersect_ensgs))
#print(intersect_ensgs)
```


==========
Definitional Biodom-GO term mappings: syn25428992 (rda)

```{r}
source <- download_file("syn25428992", "biodom_go", "rds")
assign(paste("biodom_go", "source", sep="_"), source, envir = .GlobalEnv)
```

Alpha list of Biodomain names:

```{r}
biodomains <- unique(subset(biodom_go_source, select=c("Biodomain")))
biodomains[order(biodomains$Biodomain),]
```


Mappings:
```{r}
biodom_go_source[biodom_go_source$ensembl_id %in% "ENSG00000147065",]
```


----
Investigations of gene-biodomain-goterms abbreviated csv source file (syn44151254)

```{r}
biodomains_source[biodomains_source$ensembl_id %in% "ENSG00000147065",]
```

Shiny app source RDS file (from app github repo)
TODO support loading local file via utils?
```{r}
 filename <- file.choose()
 biodom_genes <- readRDS(filename)
 biodom_genes
```

This older version of the file seems to only have 16/19 biodomains in it:
```{r}
biodoms <- unique(subset(biodom_genes, select=c("Biodomain", "bd_terms")))
biodoms[order(biodoms$Biodomain),]

```
What's in the Annotated Biological Domains (Mouse) file?

```{r}
source <- download_file("syn26592124", "mus_biodom", "rds")
assign(paste("mus_biodom", "source", sep="_"), source, envir = .GlobalEnv)
```

```{r}
mus_biodom_source
unique(subset(mus_biodom_source, select=c("Biodomain")))
```

Unique biodomains
```{r}
biodomains <- sort(unique(biodomains_source$Biodomain))
cat(biodomains, sep = "\n")
```

---
What is the longest GO term we have?
```{r}
terms <- na.omit(unique(biodomains_source$GOterm_Name))
which.max(nchar(terms))
terms[6180]
```

Find a gene that is linked to it 
-> NA
```{r}
test_term <- "immune system process"
long_term <- "oxidoreductase activity, acting on paired donors, with incorporation or reduction of molecular oxygen, reduced iron-sulfur protein as one donor, and incorporation of one atom of oxygen"
long_term_2 <- "nickel incorporation into iron-sulfur cluster via tris-L-cysteinyl L-cysteine persulfido L-glutamato L-histidino L-serinyl nickel triiron disulfide trioxide"
long_term_3 <- "iron incorporation into iron-sulfur cluster via tris-L-cysteinyl L-cysteine persulfido L-glutamato L-histidino L-serinyl nickel triiron disulfide trioxide"
long_term_4 <- "nickel incorporation into iron-sulfur cluster via tris-L-cysteinyl L-cysteine persulfido bis-L-glutamato L-histidino nickel triiron disulfide trioxide"
long_term_5 <- "endonucleolytic cleavage in ITS1 to separate SSU-rRNA from 5.8S rRNA and LSU-rRNA from tricistronic rRNA transcript (SSU-rRNA, 5.8S rRNA, LSU-rRNA)"


# row <- biodomains_source[biodomains_source$GOterm_Name %in% c(test_term),]
#row <- biodomains_source[biodomains_source$GOterm_Name %in% c(long_term),]
row <- biodomains_source[biodomains_source$GOterm_Name %in% c(long_term_5),]

row$ensembl_id
```

What is the next longest GO term we have?
```{r}
terms <- na.omit(unique(biodomains_source$GOterm_Name))
terms <- terms[-6180]
terms <- terms[-6170]
terms <- terms[-6168]
terms <- terms[-6168]
which.max(nchar(terms))
terms[which.max(nchar(terms))]
```

## AG-1245 dup ENSGs introduced via biodomains v1

```{r}
v1_source <- download_file("syn44151254.1", type = "csv") # file type: csv, tsv, json, f (feather)
assign(paste(name, "source", sep="_"), source, envir = .GlobalEnv)
```


```{r}
v3_source <- download_file("syn44151254.3", type = "csv") # file type: csv, tsv, json, f (feather)
assign(paste(name, "source", sep="_"), source, envir = .GlobalEnv)
```

```{r}
iptk <- c("ENSG00000100605", "ENSG00000274958")
iptk_rows <- v3_source[v3_source$ensembl_id %in% iptk,]
cat("\niptk: ", unique(iptk_rows$ensembl_id))

ldha <- c("ENSG00000134333", "ENSG00000288299")
ldha_rows <- v3_source[v3_source$ensembl_id %in% ldha,]
cat("\nldha: ", unique(ldha_rows$ensembl_id))

mapt <- c("ENSG00000186868", "ENSG00000277956", "ENSG00000276155")
mapt_rows <- v3_source[v3_source$ensembl_id %in% mapt,]
cat("\nmapt: ", unique(mapt_rows$ensembl_id))

ndufa6 <- c("ENSG00000184983", "ENSG00000272765", "ENSG00000277365", "ENSG00000281013", "ENSG00000273397")
ndufa6_rows <- v3_source[v3_source$ensembl_id %in% ndufa6,]
cat("\nndufa6: ", unique(ndufa6_rows$ensembl_id))

ptrpd <- c("ENSG00000153707", "ENSG00000282932")
ptrpd_rows <- v3_source[v3_source$ensembl_id %in% ptrpd,]
cat("\nptrpd: ", unique(ptrpd_rows$ensembl_id))

pvalb <- c("ENSG00000100362", "ENSG00000274665")
pvalb_rows <- v3_source[v3_source$ensembl_id %in% pvalb,]
cat("\npvalb: ", unique(pvalb_rows$ensembl_id))

ruvbl1 <- c("ENSG00000175792", "ENSG00000284901")
ruvbl1_rows <- v3_source[v3_source$ensembl_id %in% ruvbl1,]
cat("\nruvbl1: ", unique(ruvbl1_rows$ensembl_id))
```
```{r}
iptk <- c("ENSG00000100605", "ENSG00000274958")
iptk_rows <- v1_source[v1_source$ensembl_id %in% iptk,]
cat("\niptk: ", unique(iptk_rows$ensembl_id))

ldha <- c("ENSG00000134333", "ENSG00000288299")
ldha_rows <- v1_source[v1_source$ensembl_id %in% ldha,]
cat("\nldha: ", unique(ldha_rows$ensembl_id))

mapt <- c("ENSG00000186868", "ENSG00000277956", "ENSG00000276155")
mapt_rows <- v1_source[v1_source$ensembl_id %in% mapt,]
cat("\nmapt: ", unique(mapt_rows$ensembl_id))

ndufa6 <- c("ENSG00000184983", "ENSG00000272765", "ENSG00000277365", "ENSG00000281013", "ENSG00000273397")
ndufa6_rows <- v1_source[v1_source$ensembl_id %in% ndufa6,]
cat("\nndufa6: ", unique(ndufa6_rows$ensembl_id))

ptrpd <- c("ENSG00000153707", "ENSG00000282932")
ptrpd_rows <- v1_source[v1_source$ensembl_id %in% ptrpd,]
cat("\nptrpd: ", unique(ptrpd_rows$ensembl_id))

pvalb <- c("ENSG00000100362", "ENSG00000274665")
pvalb_rows <- v1_source[v1_source$ensembl_id %in% pvalb,]
cat("\npvalb: ", unique(pvalb_rows$ensembl_id))

ruvbl1 <- c("ENSG00000175792", "ENSG00000284901")
ruvbl1_rows <- v1_source[v1_source$ensembl_id %in% ruvbl1,]
cat("\nruvbl1: ", unique(ruvbl1_rows$ensembl_id))
```


```{r}
ensembl_ids <- v1_source$ensembl_id
v1_list <- grepl(";", ensembl_ids, fixed = TRUE)
v1_concats <- v1_source[v1_list,]
```

```{r}
ensembl_ids <- v3_source$ensembl_id
v3_list <- grepl(";", ensembl_ids, fixed = TRUE)
v3_concats <- v3_source[v3_list,]
v3_concats
nrow(v3_concats)

hgncs_with_v3_dups <- c("IPTK", "LDHA", "MAPT", "NDUFA6", "PTRPD", "PVALB", "RUVBL1")
v3_concats[v3_concats$symbol %in% hgncs_with_v3_dups,]
```

## AG-1306 ENSGs with null ensembl_infos introduced via biodomains source

Are these in the v3 biodomains source? -> NO
```{r}
missing_ensembl_infos <- c("ENSG00000288359", "ENSG00000288307", "ENSG00000284063", "ENSG00000283975", "ENSG00000283951", "ENSG00000283264", "ENSG00000283158")
v1_source[v1_source$ensembl_id %in% missing_ensembl_infos,]
v3_source[v3_source$ensembl_id %in% missing_ensembl_infos,]

v3_concats[is.na(v3_concats$hgnc_symbol),]
```