---
title: "metabolomics.json"
output: html_notebook
---

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```

# Validation run metadata
```{r echo=FALSE}
# file-specific
old_synId <-  "syn19315964" 
new_synId <-  "syn26868927"
name <- "metabolomics"

# not file-specific
old_name <- paste(name, "old", sep="_")
new_name <- paste(name, "new", sep="_")
rules <- paste0("rules_", name)

# print validation summary
cat( name, ".json", "\n", date(),
    "\nold: ", old_synId, " \nnew: ", new_synId,
    sep="")
```

# Downloaded files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
download_files(old_synId, new_synId, name)
```

# Compared old & new files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
compare(get(old_name), get(new_name), name, summarize=FALSE)
```


# Validated new file against rules
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_metabolomics.R")
test(get(new_name), get(rules))
```


***
# Validation failure investigations
***

## association_p > 0: 62 fails

Analysis results:  
- all fails are from association_p == 0  
- the old data file didn't have these  
- none in the source data either  

```{r}
metabolomics_new[metabolomics_new$association_p == 0, ]
```

Check for these in the old file:
```{r}
metabolomics_old[metabolomics_old$association_p == 0, ]
```


Source data (syn26064497)
```{r}
download_file("syn26064497", "metabolomics_source", type="f")
metabolomics_source_file <- synGet("syn26064497", downloadLocation = "files/")
metabolomics_source <- read_feather(metabolomics_source_file$path)
```

Check for these in the source data:
```{r}
metabolomics_df <- as.data.frame(metabolomics_source)
ass_p <- (pull(metabolomics_source, "association_p"))
length(ass_p[ass_p == 0])
```


Check specific cases in source data:
```{r}
metabolomics_source[metabolomics_source$associated_gene_name %in% c("MYH16", "AOC1", "TMEM176A", "CYP3A43", "ANAPC5", "PXN", "BUD31", "PTCD1", "CYP3A5", "ZKSCAN1"),]
```


## is_unique(associated_gene_name): 438 fails

Analysis results:  
- duplicate hgnc look similar to those identified in AG-333, fwiw...  
- source data contains dups with disjointed pvalues too  

```{r}
counts <- as.data.frame(table(metabolomics_new$associated_gene_name))
counts[order(counts$Freq, decreasing = TRUE),]

```

CRIP1 dups  

ENSG00000213145 - this is CRIP1 (https://uswest.ensembl.org/Homo_sapiens/Gene/Summary?db=core;g=ENSG00000213145;r=14:105486317-105488947)  
ENSG00000257341 - this is not CRIP1 (https://uswest.ensembl.org/Homo_sapiens/Gene/Summary?db=core;g=ENSG00000257341;r=14:105487199-105492267)  


```{r}
metabolomics_new[metabolomics_new$associated_gene_name == 'CRIP1',]
```

C17orf100 dups  

ENSG00000256806 - this ENSG is current (http://uswest.ensembl.org/Homo_sapiens/Gene/Sequence?g=ENSG00000256806;r=17:6651762-6693202)  
ENSG00000279257 - this ENSG is archived (http://dec2016.archive.ensembl.org/Homo_sapiens/Gene/Summary?db=core;g=ENSG00000279257;r=17:6651914-6652270;t=ENST00000624159)  

Why are the p-values so wildly different here?

```{r}
metabolomics_new[metabolomics_new$associated_gene_name == 'C17orf100',]
```
U3 dups  

so many ENSG for human U3...(https://uswest.ensembl.org/Human/Search/Results?q=U3;site=ensembl;facet_species=Human)  

but why are the p-vals for the same metabolite wildly different?  

```{r}
metabolomics_new[metabolomics_new$associated_gene_name == 'U3',]
```


## is_unique(associated_gene_name, metabolite_id): 231 fails

Analysis results:  
- Unsurprisingly, these overlap with the genes that aren't unique on hgnc alone (case above)  
- Some dups have different association_p values  
- Some dups have different gene_wide_threshold_1kgp values  
- Some dups have different values for both p-vals  
- Metazoa_SRP c18.2 has 3 duplicates  
- Y_RNA has 3+ dups for multiple metabolites  
- Source data spot checks find the same dups  

```{r}
new_hgnc <- metabolomics_new %>% 
  group_by(associated_gene_name, metabolite_id) %>% 
  mutate(dupe = n()>1)
dups_new_hgnc <- new_hgnc[new_hgnc$dupe == TRUE,]
dups_new_hgnc[order(dups_new_hgnc$associated_gene_name, dups_new_hgnc$metabolite_id),]
```
How many non-identical dups do we have?  
231 total dups  
- 138 dups have different association_p values  
- 207 dups have different gene_wide_p_threshold_1kgp values  
- 20 dups are identical across all four of these fields: associated_gene_name, metabolite_id, association_p, gene_wide_p_threshold_1kgp   

```{r}
new_hgnc <- metabolomics_new %>% 
  group_by(associated_gene_name, metabolite_id) %>% 
  mutate(dupe = n()>1)
dups_new_hgnc <- new_hgnc[new_hgnc$dupe == TRUE,]
dups_new_hgnc_sorted <- dups_new_hgnc[order(dups_new_hgnc$associated_gene_name, dups_new_hgnc$metabolite_id),]

new_assp <- metabolomics_new %>% 
  group_by(associated_gene_name, metabolite_id, association_p) %>% 
  mutate(dupe = n()>1)
dups_new_assp <- new_assp[new_assp$dupe == TRUE,]
dups_new_assp_sorted <- dups_new_assp[order(dups_new_assp$associated_gene_name, dups_new_assp$metabolite_id),]

new_gwpt <- metabolomics_new %>% 
  group_by(associated_gene_name, metabolite_id, gene_wide_p_threshold_1kgp) %>% 
  mutate(dupe = n()>1)
dups_new_gwpt <- new_gwpt[new_gwpt$dupe == TRUE,]
dups_new_gwpt_sorted <- dups_new_gwpt[order(dups_new_gwpt$associated_gene_name, dups_new_gwpt$metabolite_id),]

new_all <- metabolomics_new %>% 
  group_by(associated_gene_name, metabolite_id, association_p, gene_wide_p_threshold_1kgp) %>% 
  mutate(dupe = n()>1)
dups_new_all <- new_all[new_all$dupe == TRUE,]
dups_new_all_sorted <- dups_new_all[order(dups_new_all$associated_gene_name, dups_new_all$metabolite_id),]

dups_new_hgnc_sorted
setdiff(dups_new_hgnc_sorted, dups_new_assp_sorted)
setdiff(dups_new_hgnc_sorted, dups_new_gwpt_sorted)
setdiff(dups_new_hgnc_sorted, dups_new_all_sorted)
```




duplicates with differing p-values exist in the source data
```{r}
metabolomics_source[metabolomics_source$associated_gene_name == "C17orf100",]
```


```{r}
metabolomics_source[metabolomics_source$associated_gene_name == "CRIP1",]
```