---
title: "metabolomics.json"
output: html_notebook
---

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```

# Validation run metadata
```{r echo=FALSE}
# file-specific
name <- "metabolomics"
source_synId <- "syn26064497"
old_synId <-  "syn19315964"
new_synId <- "syn19276330" 

# not file-specific
source_name <- paste(name, "source", sep="_")
old_name <- paste(name, "old", sep="_")
new_name <- paste(name, "new", sep="_")
rules <- paste0("rules_", name)

# print validation summary
cat( name, ".json", "\n", date(),
    "\nold: ", old_synId, " \nnew: ", new_synId,
    sep="")
```

# Downloaded files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
download_files(old_synId, new_synId, name)
```

```{r}
source <- download_file(source_synId, type = "f")
assign(paste(name, "source", sep="_"), source, envir = .GlobalEnv)
```

# Compared old & new files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
compare(get(old_name), get(new_name), name)
```


# Validated new file against rules
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_metabolomics.R")
test(get(new_name), get(rules))
```

***
# Outstanding validation failures
***

Spot checking of the source data indicates that these duplicates are not created by our processing; note that the dups do have unique ENSGs:

* is_unique(associated_gene_name): 438 fails
* is_unique(associated_gene_name, metabolite_id): 231 fails

***
# Validation failure investigations
```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(VALIDATION_RULE)
cf <- confront(DATAFRAME, v)
out <- values(cf)
is_unique_fails <- DATAFRAME[!out,]
is_unique_fails
```


## is_unique(associated_gene_name): 438 fails

Analysis results:  
- duplicate hgnc look similar to those identified in AG-333, fwiw...  
- source data contains dups with disjointed pvalues too  


```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(is_unique(associated_gene_name))
cf <- confront(metabolomics_new, v)
out <- values(cf)
is_unique_fails <- metabolomics_new[!out,]
is_unique_fails[order(is_unique_fails$associated_gene_name),]
```


```{r}
counts <- as.data.frame(table(metabolomics_new$associated_gene_name))
dups <- counts[counts$Freq > 1,]
dups[order(dups$Freq, decreasing = TRUE),]

```

CRIP1 dups  

ENSG00000213145 - this is CRIP1 (https://uswest.ensembl.org/Homo_sapiens/Gene/Summary?db=core;g=ENSG00000213145;r=14:105486317-105488947)  
ENSG00000257341 - this is not CRIP1 (https://uswest.ensembl.org/Homo_sapiens/Gene/Summary?db=core;g=ENSG00000257341;r=14:105487199-105492267)  


```{r}
metabolomics_new[metabolomics_new$associated_gene_name == 'CRIP1',]
```

C17orf100 dups  

ENSG00000256806 - this ENSG is current (http://uswest.ensembl.org/Homo_sapiens/Gene/Sequence?g=ENSG00000256806;r=17:6651762-6693202)  
ENSG00000279257 - this ENSG is archived (http://dec2016.archive.ensembl.org/Homo_sapiens/Gene/Summary?db=core;g=ENSG00000279257;r=17:6651914-6652270;t=ENST00000624159)  

Why are the p-values so wildly different here?

```{r}
metabolomics_new[metabolomics_new$associated_gene_name == 'C17orf100',]
```
U3 dups  

so many ENSG for human U3...(https://uswest.ensembl.org/Human/Search/Results?q=U3;site=ensembl;facet_species=Human)  

but why are the p-vals for the same metabolite wildly different?  

```{r}
metabolomics_new[metabolomics_new$associated_gene_name == 'U3',]
```


## is_unique(associated_gene_name, metabolite_id): 231 fails

Analysis results:  
- Unsurprisingly, these overlap with the genes that aren't unique on hgnc alone (case above)  
- Some dups have different association_p values  
- Some dups have different gene_wide_threshold_1kgp values  
- Some dups have different values for both p-vals  
- Metazoa_SRP c18.2 has 3 duplicates  
- Y_RNA has 3+ dups for multiple metabolites  
- Source data spot checks find the same dups  

```{r}
new_hgnc <- metabolomics_new %>% 
  group_by(associated_gene_name, metabolite_id) %>% 
  mutate(dupe = n()>1)
dups_new_hgnc <- new_hgnc[new_hgnc$dupe == TRUE,]
dups_new_hgnc[order(dups_new_hgnc$associated_gene_name, dups_new_hgnc$metabolite_id),]
```
How many non-identical dups do we have?  
231 total dups  
- 138 dups have different association_p values  
- 207 dups have different gene_wide_p_threshold_1kgp values  
- 20 dups are identical across all four of these fields: associated_gene_name, metabolite_id, association_p, gene_wide_p_threshold_1kgp   

```{r}
new_hgnc <- metabolomics_new %>% 
  group_by(associated_gene_name, metabolite_id) %>% 
  mutate(dupe = n()>1)
dups_new_hgnc <- new_hgnc[new_hgnc$dupe == TRUE,]
dups_new_hgnc_sorted <- dups_new_hgnc[order(dups_new_hgnc$associated_gene_name, dups_new_hgnc$metabolite_id),]

new_assp <- metabolomics_new %>% 
  group_by(associated_gene_name, metabolite_id, association_p) %>% 
  mutate(dupe = n()>1)
dups_new_assp <- new_assp[new_assp$dupe == TRUE,]
dups_new_assp_sorted <- dups_new_assp[order(dups_new_assp$associated_gene_name, dups_new_assp$metabolite_id),]

new_gwpt <- metabolomics_new %>% 
  group_by(associated_gene_name, metabolite_id, gene_wide_p_threshold_1kgp) %>% 
  mutate(dupe = n()>1)
dups_new_gwpt <- new_gwpt[new_gwpt$dupe == TRUE,]
dups_new_gwpt_sorted <- dups_new_gwpt[order(dups_new_gwpt$associated_gene_name, dups_new_gwpt$metabolite_id),]

new_all <- metabolomics_new %>% 
  group_by(associated_gene_name, metabolite_id, association_p, gene_wide_p_threshold_1kgp) %>% 
  mutate(dupe = n()>1)
dups_new_all <- new_all[new_all$dupe == TRUE,]
dups_new_all_sorted <- dups_new_all[order(dups_new_all$associated_gene_name, dups_new_all$metabolite_id),]

dups_new_hgnc_sorted
setdiff(dups_new_hgnc_sorted, dups_new_assp_sorted)
setdiff(dups_new_hgnc_sorted, dups_new_gwpt_sorted)
setdiff(dups_new_hgnc_sorted, dups_new_all_sorted)
```


```{r}
metabolomics_source <- download_file("syn26064497", "metabolomics_source", "f")
```

duplicates with differing p-values exist in the source data
```{r}
metabolomics_source[metabolomics_source$associated_gene_name == "C17orf100",]
```


```{r}
metabolomics_source[metabolomics_source$associated_gene_name == "CRIP1",]
```

Upper case metabolite names maybe (y-axis label)?
```{r}
metabolomics_new[metabolomics_new$metabolite_full_name == "Ornithine",]
metabolomics_new$metabolite_full_name
```

How many genes do we cover?
```{r}
length(unique(metabolomics_new$ensembl_gene_id))
```