---
title: "network.json"
output: html_notebook
---

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```


# Validation run metadata
```{r echo=FALSE}
# file-specific
old_synId <-  "syn12523173" 
new_synId <-  "syn26868935"
name <- "network"

# not file-specific
old_name <- paste(name, "old", sep="_")
new_name <- paste(name, "new", sep="_")
rules <- paste0("rules_", name)

# print validation summary
cat( "File: ", name, ".json", "\nValidated on: ", date(),
    "\nOld file: ", old_synId, " \nNew file: ", new_synId,
    sep="")
```

# Downloaded files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
download_files(old_synId, new_synId, name)
```

# Compared old & new files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
compare(get(old_name), get(new_name), name, summarize=FALSE)
```


# Validated new file against rules
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_network.R")
test(get(new_name), get(rules))
```


***
# Validation failure investigations
***

## AG-: 1262 new entities
TODO


## 	is_unique(brainregion, genea_ensembl_gene_id, geneb_ensembl_gene_id, genea_external_gene_name, geneb_external_gene_name)


```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(is_unique(brainregion, genea_ensembl_gene_id, geneb_ensembl_gene_id, genea_external_gene_name, geneb_external_gene_name))
cf <- confront(network_new, v)
out <- values(cf)
is_unique_fails <- network_new[!out,]
is_unique_fails
```
```{r}
gene_info_new[gene_info_new$hgnc_symbol == 'AJ009632.2',]
gene_info_new[gene_info_new$ensembl_gene_id == 'ENSG00000229425',]
```


## field_length(genea_external_gene_name, min = 2, max = 100)
318 fails

Analysis results:  
- all 318 fails are due to blank strings  
- the old data file did not have any of these  
- Filed AG-345 for tracking (combined with next issue)  
- All missing hgncs are also missing in gene_info
=> same as AG-334



Failing cases:

```{r}
network_new[network_new$genea_external_gene_name == "", ]
```



Do we see this in the old file? -> no

```{r}
network_old[network_old$geneA_external_gene_name == "", ]
```

Spot check missing hgnc against gene_info:
```{r}
no_hgnc <- network_new[network_new$genea_external_gene_name == "", ]
no_hgnc_ensgs <- unique(no_hgnc$genea_ensembl_gene_id)

gene_info_new[gene_info_new$ensembl_gene_id %in% no_hgnc_ensgs, ]
```
Spot check missing hgnc against gene_info:
```{r}
no_hgnc <- network_new[network_new$geneb_external_gene_name == "", ]
no_hgncb_ensgs <- unique(no_hgnc$geneb_ensembl_gene_id)

gene_info_new[gene_info_new$ensembl_gene_id %in% no_hgncb_ensgs, ]
```

***
## field_length(geneb_external_gene_name, min = 2, max = 100)
216 fails

Analysis results:  
- all 216 fails are due to blank strings as gene names  
- the old data file did not have any of these  
- Filed AG-345 for tracking (combined with previous issue)  

```{r}
network_new[network_new$geneb_external_gene_name == "", ]
```



- Do we see this in the old file? -> no
```{r}
network_old[network_old$geneB_external_gene_name == "", ]
```

***

## is_unique(brainregion, genea_ensembl_gene_id, geneb_ensembl_gene_id)
144 fails

Analysis results:  
- all 144 fails are due to identical pairs of duplicate records  
- the old data file did not have any of these  
- Filed AG-346 for tracking (combined with next issue)  

```{r}
new_ensg <- network_new %>% 
  group_by(brainregion, genea_ensembl_gene_id, geneb_ensembl_gene_id) %>% 
  mutate(dupe = n()>1)
dups_new_ensg <- new_ensg[new_ensg$dupe == TRUE,]
dups_new_ensg
```


Check for these in the old file:

```{r}
old_ensg <- network_old %>% 
  group_by(brainRegion, geneA_ensembl_gene_id, geneB_ensembl_gene_id) %>% 
  mutate(dupe = n()>1)
old_ensg[old_ensg$dupe == TRUE,]
```
***
## is_unique(brainregion, genea_external_gene_name, geneb_external_gene_name)
198 fails

Analysis results:  
- all 198 fails are due to identical pairs of duplicate records  
- the old data file did not have any of these  
- Filed AG-346 for tracking (combined with previous issue)  

```{r}
new_hgnc <- network_new %>% 
  group_by(brainregion, genea_external_gene_name, geneb_external_gene_name) %>% 
  mutate(dupe = n()>1)
dups_new_hgnc <- new_hgnc[new_hgnc$dupe == TRUE,]
dups_new_hgnc[order(dups_new_hgnc$geneb_ensembl_gene_id),]
```


Check for these in the old file:

```{r}
old_hgnc <- network_old %>% 
  group_by(brainRegion, geneA_external_gene_name, geneB_external_gene_name) %>% 
  mutate(dupe = n()>1)
old_hgnc[old_hgnc$dupe == TRUE,]
```


144 records duplicate across all fields:
```{r}
new_all <- network_new %>% 
  group_by(brainregion, genea_external_gene_name, geneb_external_gene_name, genea_ensembl_gene_id, geneb_ensembl_gene_id) %>% 
  mutate(dupe = n()>1)
dups_new_all <- new_all[new_all$dupe == TRUE,]
dups_new_all
```


The 54 records that have dup ensg but not hgnc:
```{r}
dups_new_hgnc[dups_new_hgnc$geneb_external_gene_name != 'AJ009632.2', ]
```