---
title: "neuropath_corr.json"
output: html_notebook
---

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```

# Validation run metadata
```{r echo=FALSE}
# file-specific
old_synId <-  "syn22167182" 
new_synId <-  "syn26868912"
name <- "neuropath_corr"

# not file-specific
old_name <- paste(name, "old", sep="_")
new_name <- paste(name, "new", sep="_")
rules <- paste0("rules_", name)

# print validation summary
cat( name, ".json", "\n", date(),
    "\nold: ", old_synId, " \nnew: ", new_synId,
    sep="")
```

# Downloaded files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
download_files(old_synId, new_synId, name)
```

# Compared old & new files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
compare(get(old_name), get(new_name), name, summarize=FALSE)
```


# Validated new file against rules
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_neuropath_corr.R")
test(get(new_name), get(rules))
```


***
# Validation failure investigations
***
## field_length(hgnc_gene_id, min = 2, max = 100): 8516 fails

Analysis results  
* The 8516 hgnc/gnames are missing in the source data  
* We seem to have no issues displaying this data in the app - we must join on ENSG and populate the hgnc from gene_info or elsewhere  
--- see for example https://agora-develop.adknowledgeportal.org/genes/(genes-router:gene-details/ENSG00000268903)   

All of these are blank strings

```{r}
neuropath_corr_new[neuropath_corr_new$hgnc_gene_id == "",]
```
These were NAs in the old data

```{r}
nas <- is.na(neuropath_corr_old$gname)
neuropath_corr_old[nas,]
```


```{r}
neuropath_corr_source <- download_file("syn22017882", type = "csv")
nas_source <- is.na(neuropath_corr_source$GName)
neuropath_corr_source[nas_source,]
```



## RESOLVED: AG-348 - pval > 0: 19 fails

fixed via AG-349 custom json serialization fix

Analysis results:  
* All 19 have pval == 0  
* The old file and the source data do not have any pval == 0  
* Filed AG-348 for tracking

```{r}
neuropath_corr_new[neuropath_corr_new$pval == 0 ,]
```

We didn't have these in the old data

```{r}
neuropath_corr_old[neuropath_corr_old$pval == 0 ,]
```

No n/as either
```{r}
nas_old <- is.na(neuropath_corr_old$pval)
neuropath_corr_old[nas_old, ]
```


or in the source data
```{r}
neuropath_corr_source[neuropath_corr_source$PVal == 0 ,]
```

```{r}
nas_source <- is.na(neuropath_corr_source$pval)
neuropath_corr_source[nas_source, ]
```