---
title: "neuropath_corr.json"
---

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```

# Validation run metadata
```{r echo=FALSE}
# file-specific
name <- "neuropath_corr"
source_synId <- "syn22017882.5"
old_synId <-  "syn22167182" 
new_synId <-  "syn22130742" 

# not file-specific
old_name <- paste(name, "old", sep="_")
new_name <- paste(name, "new", sep="_")
rules <- paste0("rules_", name)

# print validation summary
cat( name, ".json", "\n", date(),
    "\nold: ", old_synId, " \nnew: ", new_synId,
    sep="")
```

# Downloaded files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
download_files(old_synId, new_synId, name)
```

# Compared old & new files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
compare(get(old_name), get(new_name), name)
```


# Validated new file against rules
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_neuropath_corr.R")
test(get(new_name), get(rules))
```
***
# Outstanding validation failures
***
* !is.na(gname)/field_length(gname, min = 0, max = 100): 8516 fails - missing hngc from source data

***
# Validation failure investigations
***

## field_length(hgnc_gene_id, min = 2, max = 100): 8516 fails

Analysis results  
* The 8516 hgnc/gnames are missing in the source data  
* We seem to have no issues displaying this data in the app - we must join on ENSG and populate the hgnc from gene_info or elsewhere  
--- see for example https://agora-develop.adknowledgeportal.org/genes/(genes-router:gene-details/ENSG00000268903)   

-> All of these are blank strings

Resolution: Relaxed the rule to allow blank hgnc values

```{r}
neuropath_corr_new[neuropath_corr_new$gname == "",]
```
These were NAs in the old data

```{r}
nas <- is.na(neuropath_corr_old$gname)
neuropath_corr_old[nas,]
```
These are NAs in the source data

```{r}
neuropath_corr_source <- download_file("syn22017882", type = "csv")
nas_source <- is.na(neuropath_corr_source$GName)
neuropath_corr_source[nas_source,]
```

***
# Misc investigations
***

## highest correlations?

```{r}
types <- unique(neuropath_corr_new$neuropath_type)

for (type in types) {
  rows <- neuropath_corr_new[neuropath_corr_new$neuropath_type == type,]
  max_row <- rows[which.max(rows$oddsratio),]
  cat("\nType: ", type, "Max: ", max_row$oddsratio, "ENSG: ", max_row$ensg)
}
```

```{r}

```