---
title: "overall_scores.json"
output: html_notebook
---

NOTE: ETL for this file has been disabled due to AG-699 & AG-701. Agora is pinned to Live v5 for now.

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```

# Validation run metadata
```{r echo=FALSE}
# file-specific
name <- "overall_scores"
source_synId <- "syn25575156"
old_synId <-  "syn25741025.5" # pin to m37 version
new_synId <-  "syn25740976"

# not file-specific
source_name <- paste(name, "source", sep="_")
old_name <- paste(name, "old", sep="_")
new_name <- paste(name, "new", sep="_")
rules <- paste0("rules_", name)

# print validation summary
cat( name, ".json", "\n", date(),
    "\nold: ", old_synId, " \nnew: ", new_synId,"\nsource: ", source_synId,
    sep="")
```

# Downloaded files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
download_files(old_synId, new_synId, name)
```

```{r}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source <- download_file(source_synId, type = "table")
assign(paste(name, "source", sep="_"), source, envir = .GlobalEnv)
```


# Compared old & new files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
compare(get(old_name), get(new_name), name)
```


# Validated new file against rules
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_overall_scores.R")
test(get(new_name), get(rules))
```
***
# Outstanding validation failures
***
Note that these paired failures for GeneName & for each subscore are expected (missing hgnc, missing scores):
* !is.na(GeneticsScore), in_range(GeneticsScore, min = 0, max = 3): 1340 fails -> 508 with v13 source
* !is.na(GeneName), 	field_length(GeneName, min = 0, max = 100): 5447 fails -> 5060 with v13 source
* !is.na(LiteratureScore), in_range(LiteratureScore, min = 0, max = 2): 9003 fails -> 16208 with v13 source
* !is.na(OmicsScore), in_range(OmicsScore, min = 0, max = 2): 783 fails -> 915 fails with v13 source

***
# Validation failure investigations
***

```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(RULE)
cf <- confront(overall_scores_new, v)
out <- values(cf)
is_unique_fails <- overall_scores_new[!out,]
is_unique_fails
```


***
# Other validation
***

## Cross-file validation

Compare score counts with source file & generate expected score counts used by distribution_data_test

```{r}

# total number of rows in source vs new
cat("\nTotal rows in source (syn25575156): ", nrow(overall_scores_source))
cat("\nTotal rows in new (syn26868918): ", nrow(overall_scores_new))
cat("\nDiff: ", nrow(overall_scores_source) - nrow(overall_scores_new))

# number of non-zero vs 0 overall scores
overall_source_nonzero <- overall_scores_source[overall_scores_source$Overall != 0,]
overall_source_zero <- overall_scores_source[overall_scores_source$Overall == 0,]

overall_new_nonzero <- overall_scores_new[overall_scores_new$Logsdon != 0,]
overall_new_zero <- overall_scores_new[overall_scores_new$Logsdon == 0,]
        
cat("\n\nOVERALL")                                 
cat("\noverall_scores_source$Overall != 0: ", nrow(overall_source_nonzero))
cat("\noverall_scores_source$Overall == 0: ", nrow(overall_source_zero))
cat("\noverall source (nonzero + zero)/expected: ", nrow(overall_source_nonzero) + nrow(overall_source_zero), "/", nrow(overall_scores_source))

cat("\n\noverall_scores_new$Logsdon != 0: ", nrow(overall_new_nonzero))
cat("\noverall_scores_new$Logsdon == 0: ", nrow(overall_new_zero))
cat("\noverall new (nonzero + zero)/expected: ", nrow(overall_new_nonzero) + nrow(overall_new_zero), "/", nrow(overall_scores_new))

# OMICS
genomics_source_scored <- overall_scores_source[overall_scores_source$isScored_omics == 'Y', ]
genomics_source_unscored <- overall_scores_source[overall_scores_source$isScored_omics == 'N', ]

genomics_new <- overall_scores_new$OmicsScore
genomics_new_scored <- overall_scores_new[!is.na(genomics_new),]
genomics_new_unscored <- overall_scores_new[is.na(genomics_new),]

genomics_source_zero <- overall_scores_source[overall_scores_source$OmicsScore == 0,]
genomics_new_zero <- overall_scores_new[overall_scores_new$OmicsScore == 0, ]


cat("\n\nGENOMICS")
cat("\noverall_scores_source$isScored_omics == 'Y': ", nrow(genomics_source_scored))
cat("\noverall_scores_source$isScored_omics == 'N': ", nrow(genomics_source_unscored))
cat("\nomics source (Y + N)/expected: ", nrow(genomics_source_scored) + nrow(genomics_source_unscored), "/", nrow(overall_scores_source))

cat("\n\noverall_scores_new[!is.nan(genomics_new): ", nrow(genomics_new_scored))
cat("\noverall_scores_new[is.nan(genomics_new): ", nrow(genomics_new_unscored))
cat("\nomics new (NA + !NA)/expected: ", nrow(genomics_new_scored) + nrow(genomics_new_unscored), "/", nrow(overall_scores_new))

cat("\n\noverall_scores_source$OmicsScore == 0 ", nrow(genomics_source_zero))
cat("\noverall_scores_new$OmicsScore == 0: ", nrow(genomics_new_zero))


# LIT
lit_source_scored <- overall_scores_source[overall_scores_source$isScored_lit == 'Y', ]
lit_source_unscored <- overall_scores_source[overall_scores_source$isScored_lit == 'N', ]

lit_new <- overall_scores_new$LiteratureScore
lit_new_scored <- overall_scores_new[!is.na(lit_new),]
lit_new_unscored <- overall_scores_new[is.na(lit_new),]

lit_source_zero <- overall_scores_source[overall_scores_source$LiteratureScore == 0,]
lit_new_zero <- overall_scores_new[overall_scores_new$LiteratureScore == 0, ]

cat("\n\nLIT")
cat("\noverall_scores_source$isScored_lit == 'Y': ", nrow(lit_source_scored))
cat("\noverall_scores_source$isScored_lit == 'N': ", nrow(lit_source_unscored))
cat("\nlit source (Y + N)/expected: ", nrow(lit_source_scored) + nrow(lit_source_unscored), "/", nrow(overall_scores_source))

cat("\n\noverall_scores_new[!is.na(lit_new): ", nrow(lit_new_scored))
cat("\noverall_scores_new[is.na(lit_new): ", nrow(lit_new_unscored))
cat("\nlit new (NA + !NA)/expected: ", nrow(lit_new_scored) + nrow(lit_new_unscored), "/", nrow(overall_scores_new))

cat("\n\noverall_scores_source$LiteratureScore == 0: ", nrow(lit_source_zero))
cat("\noverall_scores_new$LiteratureScore == 0: ", nrow(lit_new_zero))

# from source file, number of scored vs unscored rows for genetics
genetics_source_scored <- overall_scores_source[overall_scores_source$isScored_genetics == 'Y', ]
genetics_source_unscored <- overall_scores_source[overall_scores_source$isScored_genetics == 'N', ]

genetics_new <- overall_scores_new$GeneticsScore
genetics_new_scored <- overall_scores_new[!is.na(genetics_new),]
genetics_new_unscored <- overall_scores_new[is.na(genetics_new),]

genetics_source_zero <- overall_scores_source[overall_scores_source$GeneticsScore == 0,]
genetics_new_zero <- overall_scores_new[overall_scores_new$GeneticsScore == 0, ]

cat("\n\nGENETICS")
cat("\noverall_scores_source$isScored_genetics == 'Y': ", nrow(genetics_source_scored))
cat("\noverall_scores_source$isScored_genetics == 'N': ", nrow(genetics_source_unscored))
cat("\ngenetics source (Y + N)/expected: ", nrow(genetics_source_scored) + nrow(genetics_source_unscored), "/", nrow(overall_scores_source))

cat("\n\noverall_scores_new[!is.na(genetics_new): ", nrow(genetics_new_scored))
cat("\noverall_scores_new[is.na(genetics_new): ", nrow(genetics_new_unscored))
cat("\ngenetics new (NA + !NA)/expected: ", nrow(genetics_new_scored) + nrow(genetics_new_unscored), "/", nrow(overall_scores_new))

cat("\n\noverall_scores_source$GeneticsScore == 0: ", nrow(genetics_source_zero))
cat("\noverall_scores_new$GeneticsScore == 0: ", nrow(genetics_new_zero))

```

## v9 -> v13 validation

Are the score values projecting correctly?(source vs new) 
-> Scores are identical (via all.equal to avoid floating point number equality issues) & 0 -> NA change
```{r}
source_genetics <- source[c("ENSG", "GeneticsScore")]
source_omics <- source[c("ENSG", "OmicsScore")]
source_lit <- source[c("ENSG", "LiteratureScore")]
source_overall <- source[c("ENSG", "Overall")]

new_genetics <- overall_scores_new[c("ENSG", "GeneticsScore")]
new_omics <- overall_scores_new[c("ENSG", "OmicsScore")]
new_lit <- overall_scores_new[c("ENSG", "LiteratureScore")]
new_overall <- overall_scores_new[c("ENSG", "Logsdon")]

genetics <- merge(source_genetics, new_genetics, by = "ENSG", all = TRUE)
omics <- merge(source_omics, new_omics, by = "ENSG", all = TRUE)
lit <- merge(source_lit, new_lit, by = "ENSG", all = TRUE)
overall <- merge(source_overall, new_overall, by = "ENSG", all = TRUE)

print(genetics)
print(omics)
print(lit)
print(overall)

# unscored genes are now represented by NA instead of (score == 0 && isScored == N)
cat("Genetics scores equal: ", all.equal(genetics$GeneticsScore.x, genetics$GeneticsScore.y), "\n")
cat("Omics scores equal: ", all.equal(omics$OmicsScore.x, omics$OmicsScore.y), "\n")
cat("Lit scores equal: ", all.equal(lit$LiteratureScore.x, lit$LiteratureScore.y), "\n")
cat("Overall scores equal: ", all.equal(overall$Overall, overall$Logsdon), "\n")
```

ENSG mismatches/duplicates (old vs new)
```{r}
new_ensgs <- overall_scores_new$ENSG
old_ensgs <- overall_scores_old$ENSG

new_not_old <- setdiff(new_ensgs, old_ensgs)
old_not_new <- setdiff(old_ensgs, new_ensgs)

cat("New only: ", length(new_not_old), "\n")
cat("Old only: ", length(old_not_new), "\n")

old_dups <- anyDuplicated(old_ensgs)
new_dups <- anyDuplicated(new_ensgs)

cat("old dups: ", unique(old_ensgs[duplicated(old_ensgs) != FALSE]), sep = "\n")
cat("new dups: ", unique(new_ensgs[duplicated(new_ensgs) != FALSE]), sep = "\n")
```


HGNC NAs (new vs old)
```{r}
old_hgncs <- source[c("ENSG", "GeneName")]
new_hgnc <- overall_scores_new[c("ENSG", "GeneName")]

hgncs <- merge(old_hgncs, new_hgnc, by = "ENSG", all = TRUE)
print(hgncs)

old_blanks <- hgncs[hgncs$GeneName.x == "",]
new_nas <- hgncs[is.na(hgncs$GeneName.y),]

cat("HGNCs equal: ", all.equal(hgncs$GeneName.x, hgncs$GeneName.y), "\n")
cat("Old blanks: ", nrow(old_blanks), "\n")
cat("New NAs: ", nrow(new_nas), "\n")
```


Investigate 0 scores with isScored == N vs Y projected into JSON properly
```{r}
source_genetics_scored <- source[c("ENSG", "isScored_genetics", "GeneticsScore")]
source_omics_scored <- source[c("ENSG", "isScored_omics", "OmicsScore")]
source_lit_scored <- source[c("ENSG", "isScored_lit", "LiteratureScore")]

genetics_unscored <- source_genetics_scored[source_genetics_scored$isScored_genetics == "N",]
omics_unscored <- source_omics_scored[source_omics_scored$isScored_omics == "N",]
lit_unscored <- source_lit_scored[source_lit_scored$isScored_lit == "N",]

# all isScored == N genes have a 0 score
print(genetics_unscored)
print(omics_unscored)
print(lit_unscored)

# 0 scores for isScored == Y genes
genetics_scored <- source_genetics_scored[source_genetics_scored$GeneticsScore == 0,]
omics_scored <- source_omics_scored[source_omics_scored$OmicsScore == 0,]
lit_scored <- source_lit_scored[source_lit_scored$LiteratureScore == 0,]

genetics_zeros <- genetics_scored[genetics_scored$isScored_genetics == "Y",]
omics_zeros <- omics_scored[omics_scored$isScored_omics == "Y",]
lit_zeros <- lit_scored[lit_scored$isScored_lit == "Y",]

print(genetics_zeros)
print(omics_zeros)
print(lit_zeros)
```