---
title: "overall_scores.json"
output: html_notebook
---

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```

# Validation run metadata
```{r echo=FALSE}
# file-specific
old_synId <-  "syn25741025" 
new_synId <-  "syn26868918"
name <- "overall_scores"

# not file-specific
old_name <- paste(name, "old", sep="_")
new_name <- paste(name, "new", sep="_")
rules <- paste0("rules_", name)

# print validation summary
cat( name, ".json", "\n", date(),
    "\nold: ", old_synId, " \nnew: ", new_synId,
    sep="")
```

# Downloaded files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
download_files(old_synId, new_synId, name)
```

# Compared old & new files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
compare(get(old_name), get(new_name), name, summarize=FALSE)
```


# Validated new file against rules
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_overall_scores.R")
test(get(new_name), get(rules))
```


***
# Validation failure investigations
***

## field_length(hgnc_gene_id, min = 2, max = 100): 5745 fails

- All fails are due to blank string hgnc values  
- The old file had even more NAs  
- These are all missing lit scores, but that makes sense because gene names are used for the pubmed searches  
- Agora can display the scores with NA GeneName values just file  
- if the app doesn't already handle missing vs blank strings here, it shouldn't be too hard to fix that  


```{r}
overall_scores_new[overall_scores_new$hgnc_gene_id == "",]
```


We had even more NAs in the old file  
Note that a spot check shows that the app is able to display the NA GeneName scores:   (https://agora-develop.adknowledgeportal.org/genes/(genes-router:gene-details/ENSG00000267249))  

```{r}
overall_scores_old[is.na(overall_scores_old$GeneName),]
```



## is.numeric(literaturescore)
Analysis results:  
- lit score is a string, the other scores are still numeric  
- AG-337  


```{r}
class(overall_scores_new$literaturescore)

```