---
title: "proteomics.json"
output: html_notebook
---

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```

# Validation run metadata
```{r echo=FALSE}
# file-specific
source_synId <- "syn18689335" # csv
old_synId <-  "syn18693175" 
new_synId <- "syn18691125"  

name <- "proteomics"

# not file-specific
old_name <- paste(name, "old", sep="_")
new_name <- paste(name, "new", sep="_")
rules <- paste0("rules_", name)

# print validation summary
cat( name, ".json", "\n", date(),
    "\nold: ", old_synId, " \nnew: ", new_synId,
    sep="")
```

# Downloaded files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
download_files(old_synId, new_synId, name)
```

# Compared old & new files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
compare(get(old_name), get(new_name), name, summarize=FALSE)
```


# Validated new file against rules
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_proteomics.R")
test(get(new_name), get(rules))
```
***
# Outstanding validation failures
***
These come from the source data:

* cor_pval > 0: 7680 fails

* pval > 0: 7680 fails


***
# Validation failure investigations
***

```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(RULE)
cf <- confront(proteomics_new, v)
out <- values(cf)
is_unique_fails <- proteomics_new[!out,]
is_unique_fails
```



# field_format(uniprotid, "[OPQ][0-9][A-Z0-9]{3}[0-9]|[A-NR-Z][0-9]([A-Z][A-Z0-9]{2}[0-9]){1,2}", type = "regex"): 24 fails

We are stripping 'NA' out of uniprotId for some reason!

```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(field_format(uniprotid, "[OPQ][0-9][A-Z0-9]{3}[0-9]|[A-NR-Z][0-9]([A-Z][A-Z0-9]{2}[0-9]){1,2}", type = "regex"))
cf <- confront(proteomics_new, v)
out <- values(cf)
is_unique_fails <- proteomics_new[!out,]
is_unique_fails
```


```{r}
ensgs <- is_unique_fails$ensembl_gene_id
proteomics_old[proteomics_old$ensembl_gene_id %in% ensgs,]
```


## RESOLVED: AG-349 validation
### cor_pval > 0: 7680 fails
### pval > 0: 7680 fails

- now 7440 fails, which is the same number of 0s we had before! 
- MattF fixed via custom json serializer

```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(pval > 0)
cf <- confront(proteomics_new, v)
out <- values(cf)
is_unique_fails <- proteomics_new[!out,]
is_unique_fails
```

```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(cor_pval > 0)
cf <- confront(proteomics_new, v)
out <- values(cf)
is_unique_fails <- proteomics_new[!out,]
is_unique_fails
```

* All cor_pval fails due to cor_pval == 0 
* All pval fails due to pval == 0
* All rows with cor_pval == 0 also have pval == 0
* 36 rows have pval == 0 & cor_pval !== 0

Did the old data have this issue?
* Old file had 7680 cor_pval == NA and 7680 pval == NA

=> New file has +60 (cor_pval) and +96 (pval) unexpected 0s

Did the source data have the issue?
* Source file has 7680 cor_pval == NA and 7680 pval == NA
=> New file has +60 (cor_pval) and +96 (pval) unexpected 0s


Filed AG-349 for tracking


```{r}
cor_pval_new_zero <- proteomics_new[proteomics_new$cor_pval == 0,]
print(cor_pval_new_zero[order(cor_pval_new_zero$hgnc_gene_id, cor_pval_new_zero$tissue),])
```

```{r}
pval_new_zero <- proteomics_new[proteomics_new$pval == 0,]
print(pval_new_zero[order(pval_new_zero$hgnc_gene_id, pval_new_zero$tissue),])
```

36 have pval == 0 & cor_pval !== 0

```{r}
setdiff(pval_new_zero, cor_pval_new_zero) # 
```



```{r}
cor_pval_old_na <- proteomics_old[is.na(proteomics_old$cor_pval),]
nrow(cor_pval_old_na)
```

```{r}
pval_old_na <- proteomics_old[is.na(proteomics_old$pval),]
nrow(pval_old_na)
```

All 7680 are NA for all of the results fields in old
```{r}
cat("\nlog2_fc: ",nrow(cor_pval_old_na[is.na(cor_pval_old_na$log2_fc),])) # 7680
cat("\nci_upr: ",nrow(cor_pval_old_na[is.na(cor_pval_old_na$ci_upr),])) # 7680
cat("\nci_lwr: ",nrow(cor_pval_old_na[is.na(cor_pval_old_na$ci_lwr),])) # 7680
```


```{r}
proteomics_source <- download_file("syn18689335", "proteomics_source", type="csv")
```

```{r}
proteomics_source[proteomics_source$GeneName == 'PLEC',]
```


```{r}
cor_pval_source_na <- proteomics_source[is.na(proteomics_source$Cor_PVal),] 
nrow(cor_pval_source_na)
```


```{r}
pval_source_na <- proteomics_source[is.na(proteomics_source$PVal),]
nrow(pval_source_na)
```
```{r}
cat("\nlog2_fc: ",nrow(cor_pval_old_na[is.na(cor_pval_old_na$log2_fc),])) 
cat("\nci_upr: ",nrow(cor_pval_old_na[is.na(cor_pval_old_na$ci_upr),])) 
cat("\nci_lwr: ",nrow(cor_pval_old_na[is.na(cor_pval_old_na$ci_lwr),])) 
```


All 7680 are NA for all of the other results fields in source

```{r}
cat("\nLog2_FC: ",nrow(cor_pval_source_na[is.na(cor_pval_source_na$Log2_FC),])) 
cat("\nCI_Upr: ",nrow(cor_pval_source_na[is.na(cor_pval_source_na$CI_Upr),])) 
cat("\nCI_Lwr: ",nrow(cor_pval_source_na[is.na(cor_pval_source_na$CI_Lwr),]))
```


A few specific cases from the 36 that have pval==0 and cor_pval != 0 in new

NDUFB5|O43674 - DLPFC
- pval is populated in source

```{r}
proteomics_source[proteomics_source$UniqID == "NDUFB5|O43674",]
```

LETM1|O95202 - DLPFC
- pval is populated in source

```{r}
proteomics_source[proteomics_source$UniqID == "LETM1|O95202", ]
```

Isolating the 60 new cor_pval cases - we only see 51 here so there must be 9 dups??
```{r}
new_zero_uniqid <- cor_pval_new_zero$uniqid
source_na_uniqid <- cor_pval_source_na$UniqID
setdiff(new_zero_uniqid, source_na_uniqid)
```

Isolating the 96 new pval cases - we only see 80 here, so there must be 16 dups (9 of which should be common with the cor_pval dups)
```{r}
new_pval_zero_uniqid <- pval_new_zero$uniqid
source_pval_na_uniqid <- pval_source_na$UniqID
setdiff(new_pval_zero_uniqid, source_pval_na_uniqid)
```


Is there anything special about these cases?

```{r}
proteomics_new[proteomics_new$uniqid == "PDHX|O00330-3",]
```


```{r}
proteomics_source[proteomics_source$UniqID == "PDHX|O00330-3",]
```

```{r}
proteomics_source[proteomics_source$UniqID == "PURA|Q00577",]
```


## 	AG-338: is_unique(ensemble_gene_id, uniprotid, tissue): 24 fails


Analysis results:
* The 24 dups appear to be due to divergent hgnc values that aren't actually hgnc values but are protein names   
* These dup rows with weird hgncs are in both the old data and the source data 
* AG-338


Resolution:     
* The app currently looks up proteomics by hgnc only, so we should never get the weirdo rows   
* In some cases (e.g. ENSG00000172867), the weirdo rows (CON__P35908) are the ones with data...but again, that is how we got it from the source  
* Since only 3 ENSGs have these dups, they all come from the source data, and we don't surface the weird rows in Agora...let's do nothing right now  
* Comment out this rule until/unless we either  
** Get updated source data  
** Start querying proteomics by ENSG and have to figure out how to deal with these rows  



ENSG00000178462 == TUBAL3, there  is no TBAL3 per hgnc - TBAL is a uniprot ID: https://www.uniprot.org/uniprot/?query=TBAL3&sort=score    
ENSG00000166548 == TK2, there is no KITM per hgnc - also a uniprot ID: https://www.uniprot.org/uniprot/?query=kitm&sort=score   
ENSG00000172867 == KRT2, no CON__P35908  -- P35908 is a uniprot ID (where did CON come from?): https://www.uniprot.org/uniprot/P35908   

```{r}
v <- validator(is_unique(ensemble_gene_id, uniprotid, tissue))
cf <- confront(proteomics_new, v)
out <- values(cf)
is_unique_fails <- proteomics_new[!out,]
is_unique_fails[order(is_unique_fails$uniprotid, is_unique_fails$tissue),]
```


```{r}
proteomics_old[proteomics_old$ensembl_gene_id == "ENSG00000172867",]
```


```{r}
proteomics_source[proteomics_source$ENSG == "ENSG00000178462",]
```

```{r}
proteomics_source[proteomics_source$ENSG == "ENSG00000166548",]
```


```{r}
proteomics_source[proteomics_source$ENSG == "ENSG00000172867",]
```



```{r}
f <- proteomics_source[is.na(proteomics_source$Log2_FC),]
p <- proteomics_source[is.na(proteomics_source$Cor_PVal),]
u <- proteomics_source[is.na(proteomics_source$CI_Upr),]
l <- proteomics_source[is.na(proteomics_source$CI_Lwr),]

fe <- f$ENSG
pe <- p$ENSG
ue <- u$ENSG
le <- l$ENSG



setdiff(fe, pe)
setdiff(fe, ue)
setdiff(fe, le)
```



```{r}
sum(is.na(proteomics_source$Cor_PVal))
```

```{r}
sum(is.na(proteomics_source$Log2_FC))
```

```{r}
sum(is.na(proteomics_source$CI_Upr))
```

```{r}
sum(is.na(proteomics_source$CI_Lwr))
```