---
title: "proteomics_tmt.json"
---

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```

# Validation run metadata
```{r echo=FALSE}
# file-specific
name <- "proteomics_tmt"
source_synId <- "syn35221005.2" # csv
old_synId <-  "syn33344657" 
new_synId <- "syn32210527"

# not file-specific
old_name <- paste(name, "old", sep="_")
new_name <- paste(name, "new", sep="_")
rules <- paste0("rules_", name)

# print validation summary
cat( name, ".json", "\n", date(),
    "\nold: ", old_synId, " \nnew: ", new_synId,
    sep="")
```

# Downloaded files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
download_files(old_synId, new_synId, name)
```

```{r}
source <- download_file(source_synId, type = "csv")
assign(paste(name, "source", sep="_"), source, envir = .GlobalEnv)
```


# Compared old & new files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
compare(get(old_name), get(new_name), name)
```


# Validated new file against rules
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_proteomics_tmt.R")
test(get(new_name), get(rules))
```
***
# Outstanding validation failures
***
N/A

***
# Validation failure investigations
***

```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(RULE)
cf <- confront(proteomics_new, v)
out <- values(cf)
is_unique_fails <- proteomics_new[!out,]
is_unique_fails
```


***
Other validation
***
## tmt/lfq conflict
Jake's file: -4.188784	l2fc
Laura's file: 0.2077565 l2fc -> much better agreement now!
LFQ: 0.2077565 l2fc
```{r}
lfq <- proteomics_new[proteomics_new$uniqid == 'MSN|P26038',]
tmt <- proteomics_tmt_new[proteomics_tmt_new$uniqid == 'MSN|P26038',]
lfq_dlpfc <- lfq[lfq$tissue == 'DLPFC',]
tmt
lfq_dlpfc
```

## Jake vs Laura results
compare string differences
```{r}
ensg_new <- proteomics_tmt_new$ensembl_gene_id
ensg_old <- proteomics_tmt_old$ensembl_gene_id
setdiff(ensg_new, ensg_old)
```
```{r}
uniqid_new <- proteomics_tmt_new$uniqid
uniqid_old <- proteomics_tmt_old$uniqid
setdiff(uniqid_new, uniqid_old)
```
```{r}
uniprotid_new <- proteomics_tmt_new$uniprotid
uniprotid_old <- proteomics_tmt_old$uniprotid
setdiff(uniprotid_new, uniprotid_old)
```


## Boolean flip investigation
YARS|P54577 should flip when we incorporate TMT evidence:

```{r}
proteomics_new[proteomics_new$ensembl_gene_id == 'ENSG00000134684',] # no significant results
proteomics_tmt_new[proteomics_tmt_new$ensembl_gene_id == 'ENSG00000134684',] # significant result
```
TBK1|Q9UHD2 should not flip
```{r}
proteomics_new[proteomics_new$ensembl_gene_id == 'ENSG00000183735',] # no significant results
proteomics_tmt_new[proteomics_tmt_new$ensembl_gene_id == 'ENSG00000183735',] # no significant result
```

