---
title: "proteomics.json"
---

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```

# Validation run metadata
```{r echo=FALSE}
# file-specific
name <- "proteomics"
source_synId <- "syn18689335" # csv
old_synId <-  "syn18693175" # m46 v12 Live
new_synId <- "syn18691125"# v55 Testing

# not file-specific
old_name <- paste(name, "old", sep="_")
new_name <- paste(name, "new", sep="_")
rules <- paste0("rules_", name)

# print validation summary
cat( name, ".json", "\n", date(),
    "\nold: ", old_synId, " \nnew: ", new_synId,
    sep="")
```

# Downloaded files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
download_files(old_synId, new_synId, name)
```

```{r}
source <- download_file(source_synId, type = "csv")
assign(paste(name, "source", sep="_"), source, envir = .GlobalEnv)
```



# Compared old & new files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
compare(get(old_name), get(new_name), name)
```


# Validated new file against rules
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_proteomics.R")
test(get(new_name), get(rules))
```
***
# Outstanding validation failures
***
These come from the source data:

* cor_pval > 0: 7680 fails

* pval > 0: 7680 fails


***
# Validation failure investigations
***

```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(RULE)
cf <- confront(proteomics_new, v)
out <- values(cf)
is_unique_fails <- proteomics_new[!out,]
is_unique_fails
```


### cor_pval > 0: 7680 fails
### pval > 0: 7680 fails

These are all due to NA values

```{r}
proteomics_new[is.na(proteomics_new$cor_pval),]
proteomics_new[is.na(proteomics_new$pval),]
```

Source file

```{r}
cor_pval_source_na <- proteomics_source[is.na(proteomics_source$Cor_PVal),] 
nrow(cor_pval_source_na)
```

```{r}
pval_source_na <- proteomics_source[is.na(proteomics_source$PVal),] 
nrow(pval_source_na)
```

All 7680 are NA for all of the other results fields in source

```{r}
cat("\nLog2_FC: ",nrow(cor_pval_source_na[is.na(cor_pval_source_na$Log2_FC),])) 
cat("\nCI_Upr: ",nrow(cor_pval_source_na[is.na(cor_pval_source_na$CI_Upr),])) 
cat("\nCI_Lwr: ",nrow(cor_pval_source_na[is.na(cor_pval_source_na$CI_Lwr),]))
```


## 	AG-338: is_unique(ensemble_gene_id, uniprotid, tissue): 24 fails


Analysis results:
* The 24 dups appear to be due to divergent hgnc values that aren't actually hgnc values but are protein names   
* These dup rows with weird hgncs are in both the old data and the source data 
* AG-338


Resolution:     
* The app currently looks up proteomics by hgnc only, so we should never get the weirdo rows -> no longer true though
* In some cases (e.g. ENSG00000172867), the weirdo rows (CON__P35908) are the ones with data...but again, that is how we got it from the source  
* Since only 3 ENSGs have these dups, they all come from the source data, and we don't surface the weird rows in Agora...let's do nothing right now  
* Comment out this rule until/unless we either  
** Get updated source data  
** Start querying proteomics by ENSG and have to figure out how to deal with these rows  



ENSG00000178462 == TUBAL3, there  is no TBAL3 per hgnc - TBAL is a uniprot ID: https://www.uniprot.org/uniprot/?query=TBAL3&sort=score    
ENSG00000166548 == TK2, there is no KITM per hgnc - also a uniprot ID: https://www.uniprot.org/uniprot/?query=kitm&sort=score   
ENSG00000172867 == KRT2, no CON__P35908  -- P35908 is a uniprot ID (where did CON come from?): https://www.uniprot.org/uniprot/P35908   

```{r}
v <- validator(is_unique(ensemble_gene_id, uniprotid, tissue))
cf <- confront(proteomics_new, v)
out <- values(cf)
is_unique_fails <- proteomics_new[!out,]
is_unique_fails[order(is_unique_fails$uniprotid, is_unique_fails$tissue),]
```


```{r}
proteomics_old[proteomics_old$ensembl_gene_id == "ENSG00000172867",]
```


```{r}
proteomics_source[proteomics_source$ENSG == "ENSG00000178462",]
```

```{r}
proteomics_source[proteomics_source$ENSG == "ENSG00000166548",]
```


```{r}
proteomics_source[proteomics_source$ENSG == "ENSG00000172867",]
```



```{r}
f <- proteomics_source[is.na(proteomics_source$Log2_FC),]
p <- proteomics_source[is.na(proteomics_source$Cor_PVal),]
u <- proteomics_source[is.na(proteomics_source$CI_Upr),]
l <- proteomics_source[is.na(proteomics_source$CI_Lwr),]

fe <- f$ENSG
pe <- p$ENSG
ue <- u$ENSG
le <- l$ENSG



setdiff(fe, pe)
setdiff(fe, ue)
setdiff(fe, le)
```



```{r}
sum(is.na(proteomics_source$Cor_PVal))
```

```{r}
sum(is.na(proteomics_source$Log2_FC))
```

```{r}
sum(is.na(proteomics_source$CI_Upr))
```

```{r}
sum(is.na(proteomics_source$CI_Lwr))
```


***
# Other Validation
***

# l2fc characteristics (opacity and diameter tuning)
```{r}
l2fc <- proteomics_new$log2_fc
max <- max(l2fc)
min <- min(l2fc)
sd <- sd(l2fc)
cat(max, min, sd)
```

# l2fc and pval distributions (opacity and diameter tuning)

```{r}
l2fc <- proteomics_new$log2_fc
negatives <- l2fc[l2fc < 0]
positives <- l2fc[l2fc > 0]
sd <- sd(l2fc)
#sd <- 0.1467328 #rna sd
smalls <- l2fc[l2fc < -sd]
bigs <- l2fc[l2fc > sd]
small_range <- smalls[smalls > -(sd*2)]
big_range <- bigs[bigs < (sd*2)]
hist(l2fc)
hist(smalls)
hist(bigs)
hist(small_range)
hist(big_range)
cat("total negative:", length(negatives), "total positive:", length(positives), "interesting negatives", length(small_range), "interesting positives", length(big_range))
```
tuning
tuning
```{r}
l2fc <- proteomics_new$log2_fc
negatives <- l2fc[l2fc < 0]
positives <- l2fc[l2fc > 0]

low <- 0.1
high <- 0.5

purple_range <- l2fc[l2fc >= -high]
purple_range <- purple_range[purple_range <= -low]

# in range
length(purple_range)
length(purple_range)/length(negatives)
#min(purple_range)
#max(purple_range)

# out of range
high_purples <- negatives[negatives < -high]
low_purples <- negatives[negatives > -low]
length(high_purples)
length(high_purples)/length(negatives)
length(low_purples)
length(low_purples)/length(negatives)

green_range <- l2fc[l2fc >= low]
green_range <- green_range[green_range <= high]

# in range
length(green_range)
length(green_range)/length(positives)
#min(green_range)
#max(green_range)

# out of range
high_greens <- positives[positives > high]
low_greens <- positives[positives < low]
length(high_greens)
length(high_greens)/length(positives)
length(low_greens)
length(low_greens)/length(positives)
```


use 1.5IQR to eliminate outliers?
```{r}
l2fc <- rna_seq_differential_expression_new$logfc
median <- median(l2fc)
q1 <- quantile(l2fc, c(0.25), names = FALSE)
q3 <- quantile(l2fc, c(0.75), names = FALSE)
IQR <- q3 - q1
max <- q3 + (1.5 * IQR)
min <- q1 - (1.5 * IQR)

cat(min, q1, median, q3, max)

```


Boolean flip investigation
```{r}
proteomics_new[proteomics_new$ensembl_gene_id == 'ENSG00000134684',]
```


How many genes/proteins do we cover?
```{r}
length(unique(proteomics_new$ensembl_gene_id))
length(unique(proteomics_new$uniprotid))
```