---
title: "proteomics_srm.json"
---

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```

# Validation run metadata
```{r echo=FALSE}
# file-specific
name <- "proteomics_srm"
source_synId <- "syn52579640.3" # csv
old_synId <-  "syn52913548" 
new_synId <- "syn52579910"

# not file-specific
old_name <- paste(name, "old", sep="_")
new_name <- paste(name, "new", sep="_")
rules <- paste0("rules_", name)

# print validation summary
cat( name, ".json", "\n", date(),
    "\nold: ", old_synId, " \nnew: ", new_synId,
    sep="")
```

# Downloaded files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
download_files(old_synId, new_synId, name)
```

```{r}
source <- download_file(source_synId, type = "csv")
assign(paste(name, "source", sep="_"), source, envir = .GlobalEnv)
```


# Compared old & new files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
compare(get(old_name), get(new_name), name)
```


# Validated new file against rules
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_proteomics_srm.R")
test(get(new_name), get(rules))
```
***
# Outstanding validation failures
***
	pval > 0
	3 fails

	cor_pval > 0
	3 fails
	
	is_unique(hgnc_symbol, uniprotid, tissue)
	18 fails	
	
***
# Validation failure investigations
***
0 pval/cor_pval
per Jaclyn, these pvals are so small that the code rounds them to 0....see AG-1227 for notes

```{r}
proteomics_srm_source[proteomics_srm_source$ENSG == "ENSG00000115461",]
proteomics_srm_source[proteomics_srm_source$ENSG == "ENSG00000106089",]
proteomics_srm_source[proteomics_srm_source$ENSG == "ENSG00000128564",]
```


```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(pval > 0)
cf <- confront(proteomics_srm_new, v)
out <- values(cf)
is_unique_fails <- proteomics_srm_new[!out,]
is_unique_fails
```
```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(cor_pval > 0)
cf <- confront(proteomics_srm_new, v)
out <- values(cf)
is_unique_fails <- proteomics_srm_new[!out,]
is_unique_fails
```

```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(is_unique(uniprotid, tissue))
cf <- confront(proteomics_srm_new, v)
out <- values(cf)
is_unique_fails <- proteomics_srm_new[!out,]
is_unique_fails
```

***
Other validation
***
SRM tissues
```{r}
unique(proteomics_srm_new$tissue)

```

