---
title: "rna_seq_differential_expression.json"
---

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```

# Validation run metadata
```{r echo=FALSE}
# file-specific
source_synId <- "syn27211942.1" # tsv
old_synId <-  "syn12177499" 
new_synId <- "syn17015360" 

name <- "rna_seq_differential_expression"

# not file-specific
old_name <- paste(name, "old", sep="_")
new_name <- paste(name, "new", sep="_")
rules <- paste0("rules_", name)

# print validation summary
cat( name, ".json", "\n", date(),
    "\nold: ", old_synId, " \nnew: ", new_synId,
    sep="")
```

# Download files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
download_files(old_synId, new_synId, name)
```

```{r}
source <- download_file(source_synId, type = "tsv") 
assign(paste(name, "source", sep="_"), source, envir = .GlobalEnv)
```

# Compared old & new files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
compare(get(old_name), get(new_name), name)
```


# Validated new file against rules
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_rna_seq_differential_expression.R")
test(get(new_name), get(rules))
```

***
# Outstanding validation failures
***

N/A

***
# Validation failure investigations
***

```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(RULE)
cf <- confront(rna_seq_differential_expression_new, v)
out <- values(cf)
is_unique_fails <- rna_seq_differential_expression_new[!out,]
is_unique_fails
```


## is_unique(model, tissue, ensembl_gene_id): 72 fails (AG-547)

These are due to pairs of identical duplicate records for every model/tissue combination fro a single gene: 
9 x 4 = 36 expected x 2 for duplicates = 72

Not sure why we have dups, but they should be harmless.


```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(	is_unique(model, tissue, ensembl_gene_id))
cf <- confront(rna_seq_differential_expression_new, v)
out <- values(cf)
is_unique_fails <- rna_seq_differential_expression_new[!out,]
is_unique_fails
```

Are the dups from the source table?
36 matching rows in source,which equals expected rows
=> no

```{r}
rows <- rna_seq_differential_expression_source[rna_seq_differential_expression_source$ensembl_gene_id == 'ENSG00000229425',]

mayo <- rows[rows$Study == "MAYO",] # 2 regions: 2 x 4 = 8 rows expected in source
mssm <- rows[rows$Study == "MSSM",] # 4 regions: 4 x 4 = 16 rows expected in source
rosmap <- rows[rows$Study == "ROSMAP",] # 3 regions: 3 x 4 = 12 rows expected in source

print(mayo)
print(mssm)
print(rosmap)
```


***
# Other Validation
***

# 3.3.0 setdiff
```{r}
setdiff(rna_seq_differential_expression_new, rna_seq_differential_expression_old)
setdiff(rna_seq_differential_expression_old, rna_seq_differential_expression_new)

gene_new <- rna_seq_differential_expression_new[rna_seq_differential_expression_new$ensembl_gene_id =="ENSG00000163221",]
gene_old <- rna_seq_differential_expression_old[rna_seq_differential_expression_old$ensembl_gene_id =="ENSG00000163221",]

row_new <- gene_new[gene_new$tissue=="CBE",]
row_old <- gene_old[gene_old$tissue=="CBE",]

gene_join <- merge(gene_old, gene_new, by=c("ensembl_gene_id", "tissue", "model"), all = TRUE)

gene_new$match <- Reduce(`&`, Map(`==`, gene_new, gene_old))

gene_new
gene_join
```

# l2fc and pval distributions (opacity and diameter tuning)

```{r}
l2fc <- rna_seq_differential_expression_new$logfc
smalls <- l2fc[l2fc < -0.8]
bigs <- l2fc[l2fc > 0.8]
hist(smalls)
hist(bigs)
```


use 1.5IQR to eliminate outliers?
```{r}
l2fc <- rna_seq_differential_expression_new$logfc
median <- median(l2fc)
q1 <- quantile(l2fc, c(0.25), names = FALSE)
q3 <- quantile(l2fc, c(0.75), names = FALSE)
IQR <- q3 - q1
max <- q3 + (1.5 * IQR)
min <- q1 - (1.5 * IQR)

cat(min, q1, median, q3, max)

```

# unique genes we have results for
```{r}
ensgs <- rna_seq_differential_expression_new$ensembl_gene_id
length(unique(ensgs))
nrow(rna_seq_differential_expression_new)

ensgs <- rna_seq_differential_expression_old$ensembl_gene_id
length(unique(ensgs))
nrow(rna_seq_differential_expression_old)

ensgs <- rna_seq_differential_expression_source$ensembl_gene_id
length(unique(ensgs))
nrow(rna_seq_differential_expression_source)
```

TEMP
old vs new source file: included models
```{r}
source <- download_file("syn14237651", type = "tsv") 
assign(paste(name, "old_source", sep="_"), source, envir = .GlobalEnv)
```

```{r}
unique(rna_seq_differential_expression_source$Model)
unique(rna_seq_differential_expression_old_source$Model)
```






poster stats
```{r}
nrow(rna_seq_differential_expression_new)
length(unique((rna_seq_differential_expression_new$ensembl_gene_id)))
```