---
title: "rna_seq_differential_expression.json"
output: html_notebook
---

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```

# Validation run metadata
```{r echo=FALSE}
# file-specific
old_synId <-  "syn12177499" 
new_synId <-  "syn26868926"
name <- "rna_seq_differential_expression"

# not file-specific
old_name <- paste(name, "old", sep="_")
new_name <- paste(name, "new", sep="_")
rules <- paste0("rules_", name)

# print validation summary
cat( name, ".json", "\n", date(),
    "\nold: ", old_synId, " \nnew: ", new_synId,
    sep="")
```

# Downloaded files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
download_files(old_synId, new_synId, name)
```

# Compared old & new files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
compare(get(old_name), get(new_name), name, summarize=FALSE)
```


# Validated new file against rules
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_rna_seq_differential_expression.R")
test(get(new_name), get(rules))
```


***
# Validation failure investigations
***

## info for MattC

```{r}
not0 <- rna_seq_differential_expression_new[rna_seq_differential_expression_new$adj_p_val >0,]

max_lfc <- max(rna_seq_differential_expression_new$logfc)
min_lfc <- min(rna_seq_differential_expression_new$logfc)
max_adjP <- max(rna_seq_differential_expression_new$adj_p_val)
min_adjP <- min(not0$adj_p_val)

cat("\nLFC \n  max: ", max_lfc, "\n  min: ", min_lfc, "\nadj_p_val \n  max: ", max_adjP, "\n  min: ", min_adjP)
```
```{r}

clean <- rna_seq_differential_expression_old[!is.na(rna_seq_differential_expression_old$logfc),]

max_lfc <- max(clean$logfc)
min_lfc <- min(clean$logfc)
max_adjP <- max(clean$adj_p_val)
min_adjP <- min(clean$adj_p_val)

cat("\nLFC \n  max: ", max_lfc, "\n  min: ", min_lfc, "\nadj_p_val \n  max: ", max_adjP, "\n  min: ", min_adjP)
```


## AG-369: 1092 extra entities came from where?

TODO


```{r}

```
```{r}
sqldf('SELECT * from rna_seq_differential_expression_new where ensembl_gene_id ="ENSG00000000003" and tissue = "FP"')
```

```{r}
sqldf('SELECT * from rna_seq_differential_expression_old where ensembl_gene_id ="ENSG00000000003" and tissue = "FP"')
```


## 	field_length(ensembl_gene_id, n = 15): 1 fail


```{r}
v <- validator(field_length(ensembl_gene_id, n = 15))
cf <- confront(get(new_name), v)
out <- values(cf)
is_unique_fails <- get(new_name)[!out,]
is_unique_fails
```


## field_length(hgnc_symbol, min = 2, max = 100): 40400 fails

- All of these are due to missing hgnc
- 2342 ENSGs / expect 4 per region for the 7 models so up to 28 dups per ENSG is expected
- but some ENSG  have 56 dups
=> dups appear identical, so generally harmless?

```{r}
v <- validator(field_length(hgnc_symbol, min = 2, max = 100))
cf <- confront(rna_seq_differential_expression_new, v)
out <- values(cf)
is_unique_fails <- rna_seq_differential_expression_new[!out,]
print(is_unique_fails[order(is_unique_fails$ensembl_gene_id, is_unique_fails$tissue, is_unique_fails$model),])
dup_ensg <- as.data.frame(table(is_unique_fails$ensembl_gene_id))
dup_ensg[order(-dup_ensg$Freq),]

```

Rows for the 56 dups case seem identical pairwise, so 28 uniques as expected

```{r}
dups_56 <- rna_seq_differential_expression_new[rna_seq_differential_expression_new$ensembl_gene_id == "ENSG00000229425",]
dups_56[order(dups_56$tissue, dups_56$model),]
```



## is_unique(model, tissue, ensembl_gene_id): 168 fails

- We have identical dups => harmless


```{r}
v <- validator(is_unique(model, tissue, ensembl_gene_id))
cf <- confront(rna_seq_differential_expression_new, v)
out <- values(cf)
is_unique_fails <- rna_seq_differential_expression_new[!out,]
is_unique_fails[order(is_unique_fails$ensembl_gene_id, is_unique_fails$tissue, is_unique_fails$model),]
```


## 	adj_p_val > 0: 100 fails

Analysis results  
- The old data file had <2k of these  
- The source data has only 18 NAs  
- AG-350  

```{r}
v <- validator(adj_p_val > 0)
cf <- confront(rna_seq_differential_expression_new, v)
out <- values(cf)
is_unique_fails <- rna_seq_differential_expression_new[!out,]
is_unique_fails[order(is_unique_fails$ensembl_gene_id, is_unique_fails$tissue),]
```
old file has more of these (>2k)

```{r}
rna_seq_differential_expression_old[rna_seq_differential_expression_old$adj_p_val == 0,]
```

Get the source file
```{r}
rna_seq_differential_expression_source <- download_file('syn14237651', rna_seq_differential_expression_source, type = 'tsv')
```

look at a specific ENSG from the 100 fails:  
ENSG00000003989 - we have a pvals in the source data   
=> where did they go during processing?  

```{r}
ensg1 <- rna_seq_differential_expression_source[rna_seq_differential_expression_source$ensembl_gene_id == 'ENSG00000003989',]
ensg1_filtered <- ensg1[ensg1$Tissue == 'TCX',]
print(ensg1_filtered)

ensg1_more_filtered <- ensg1_filtered[ensg1_filtered$Comparison == "AD-CONTROL",]
print(ensg1_more_filtered)

diagnoses <- c("Diagnosis", "Diagnosis.AOD", "Diagnosis.Sex")
ensg1_extra_filtered <- ensg1_more_filtered[ensg1_more_filtered$Model %in% diagnoses, ]
ensg1_extra_filtered
```



source file has 1 actual 0 and 92 empty rows(I think?)  
source file has 92 NAs  
```{r}
source_zero <- rna_seq_differential_expression_source[rna_seq_differential_expression_source[[11]] == 0,]
source_na <- rna_seq_differential_expression_source[is.na(rna_seq_differential_expression_source[[11]]),]
print(source_zero)
print(source_na)
```



ENSG00000229807 - has the only row with 0 for pvals in the source data    
- The source data contains one duplicate row for DLPFC AD-CONTROL ALL (Tissue Comparison Sex)    
- The dup rows have different values for the numerics except AveExpr  
- The dup is not the row that has 0 pvals, the zero pvals row is not processed due to Comparison filtering  

```{r}
new_xist <- rna_seq_differential_expression_new[rna_seq_differential_expression_new$ensembl_gene_id == 'ENSG00000229807',]
print(new_xist[order(new_xist$tissue, new_xist$model),])

source_xist <- rna_seq_differential_expression_source[rna_seq_differential_expression_source$ensembl_gene_id == 'ENSG00000229807',]
print(source_xist)

diagnoses <- c("Diagnosis", "Diagnosis.AOD", "Diagnosis.Sex")
diagnoses_filtered_xist <- source_xist[source_xist$Model %in% diagnoses,]
print(diagnoses_filtered_xist)

double_filtered_xist <- diagnoses_filtered_xist[diagnoses_filtered_xist$Comparison == "AD-CONTROL",]

double_filtered_xist[order(double_filtered_xist$Tissue, double_filtered_xist$Model),]
```

Do the 100 new 0s include the 92 NAs? - not entirely overlapping  

-> not sure yet...  


```{r}
source_na_ensg <- source_na$ensembl_gene_id
new_zero <- rna_seq_differential_expression_new[rna_seq_differential_expression_new$adj_p_val == 0,]
new_zero_ensg <- new_zero$ensembl_gene_id


ensgs <- setdiff(new_zero_ensg, source_na_ensg)
cat(sort(ensgs), sep = "\n")

diff <- rna_seq_differential_expression_new[rna_seq_differential_expression_new$ensembl_gene_id %in% ensgs,]
filtered_diff <- diff[diff$adj_p_val == 0,]
print(filtered_diff[order(filtered_diff$ensembl_gene_id, filtered_diff$tissue, filtered_diff$hgnc_symbol),]) # in new not source

reverse_ensgs <- setdiff(source_na_ensg, new_zero_ensg)
cat(sort(reverse_ensgs), sep = "\n")

reverse_diff <-rna_seq_differential_expression_new[rna_seq_differential_expression_new$ensembl_gene_id %in% reverse_ensgs,]
print(reverse_diff[order(reverse_diff$ensembl_gene_id, reverse_diff$tissue, reverse_diff$hgnc_symbol),]) # in source not new
```


