---
title: "rna_seq_differential_expression.json"
output: html_notebook
---

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```

# Validation run metadata
```{r echo=FALSE}
# file-specific
old_synId <-  "syn12177499" 
# new_synId <-  "syn26868926" # test_data_full
new_synId <- "syn27408149" # test_data_release_1.0.0

name <- "rna_seq_differential_expression"

# not file-specific
old_name <- paste(name, "old", sep="_")
new_name <- paste(name, "new", sep="_")
rules <- paste0("rules_", name)

# print validation summary
cat( name, ".json", "\n", date(),
    "\nold: ", old_synId, " \nnew: ", new_synId,
    sep="")
```

# Downloaded files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
download_files(old_synId, new_synId, name)
```

# Compared old & new files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
compare(get(old_name), get(new_name), name, summarize=FALSE)
```


# Validated new file against rules
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
source("~/repos/rstudio/agora-data-validation/src/rules/rules_rna_seq_differential_expression.R")
test(get(new_name), get(rules))
```

***
# Outstanding validation failures
***

* is_unique(model, tissue, ensembl_gene_id): 168 fails => these duplicates should not cause issues in the app

* adj_p_val > 0: 32 fails - See AG-350

***
# Validation failure investigations
***

## model %in% models: 345027 fails

models <- c("AD Diagnosis (males and females)", "AD Diagnosis x AOD (males and females)","AD Diagnosis x Sex (females only)", "AD Diagnosis x Sex (males only)") : 345027 fails

* We are missing the "AD" from 3/4 model values -> AG-410

```{r}
# yields a dataframe containing the failure cases for a specific rule
v <- validator(model %in% c("AD Diagnosis (males and females)", "AD Diagnosis x AOD (males and females)","AD Diagnosis x Sex (females only)", "AD Diagnosis x Sex (males only)"))
cf <- confront(rna_seq_differential_expression_new, v)
out <- values(cf)
is_unique_fails <- rna_seq_differential_expression_new[!out,]
is_unique_fails
```

```{r}
unique(is_unique_fails[c("model")])
```


## is_unique(model, tissue, ensembl_gene_id): 168 fails

These fail because we have multiple hgncs mapped to the same ensg, e.g. MIR29A & MIR29B1 both map to ENSG00000226380
-> belong to the same 'gene group' of microRNAs: https://www.genenames.org/data/genegroup/#!/group/1701

Since we will do lookups by ensg, we could get either of the duplicate records.
luckily, the duplicate records appear to be identical other than hgnc, so this should not be a problem


```{r}
v <- validator(is_unique(model, tissue, ensembl_gene_id))
cf <- confront(rna_seq_differential_expression_new, v)
out <- values(cf)
is_unique_fails <- rna_seq_differential_expression_new[!out,]
is_unique_fails[order(is_unique_fails$ensembl_gene_id, is_unique_fails$tissue, is_unique_fails$model),]
```


## 	adj_p_val > 0: 32 fails

Analysis results  
- The source data has only 18 NAs  
- All fails map to "ENSG00000050628" or "ENSG00000127903"
- All fails map to the MSSM (source) / MSBB (json) study
- AG-350  


```{r}
v <- validator(adj_p_val > 0)
cf <- confront(rna_seq_differential_expression_new, v)
out <- values(cf)
is_unique_fails <- rna_seq_differential_expression_new[!out,]
is_unique_fails[order(is_unique_fails$ensembl_gene_id, is_unique_fails$tissue),]
```
-> all fails map to two ENSGs: ENSG00000050628, ENSG00000127903

old file has more of these (>2k)

```{r}
rna_seq_differential_expression_old[rna_seq_differential_expression_old$adj_p_val == 0,]
```

Get the source file
```{r}
rna_seq_differential_expression_source <- download_file('syn14237651', rna_seq_differential_expression_source, type = 'tsv')
```

Compare source vs new rows for the two ENSGs
- source file has 32 rows with no values of interest, all seem to belong to the MSSM study
- new file has 32 rows with no values of interest, all seem to belong to the MSSB study

```{r}
rna_seq_differential_expression_source[is.na(rna_seq_differential_expression_source$adj.P.Val),]

ensgs <- unique(is_unique_fails$ensembl_gene_id)
source_rows <- rna_seq_differential_expression_source[rna_seq_differential_expression_source$ensembl_gene_id %in% ensgs,]

diagnoses <- c("Diagnosis", "Diagnosis.AOD", "Diagnosis.Sex")
source_f1 <- source_rows[source_rows$Comparison == "AD-CONTROL",]
source_f2 <- source_f1[source_f1$Model %in% diagnoses, ]

source_f2[order(source_f2$Tissue, source_f2$Model),]
```
```{r}
ensgs <- unique(is_unique_fails$ensembl_gene_id)

new_rows <- rna_seq_differential_expression_new[rna_seq_differential_expression_new$ensembl_gene_id %in% ensgs,]
new_rows[order(new_rows$tissue, new_rows$model),]
```

Do we have good data for other genes from that study?

```{r}
rna_seq_differential_expression_source[rna_seq_differential_expression_source$Study == 'MSSM',]
```





look at a specific ENSG from the fails:  
ENSG00000050628 - we have a pvals in the source data   
=> where did they go during processing?  

```{r}
ensg1 <- rna_seq_differential_expression_source[rna_seq_differential_expression_source$ensembl_gene_id == 'ENSG00000050628',]
ensg1_filtered <- ensg1[ensg1$Tissue == 'TCX',]
print(ensg1_filtered)

ensg1_more_filtered <- ensg1_filtered[ensg1_filtered$Comparison == "AD-CONTROL",]
print(ensg1_more_filtered)

diagnoses <- c("Diagnosis", "Diagnosis.AOD", "Diagnosis.Sex")
ensg1_extra_filtered <- ensg1_more_filtered[ensg1_more_filtered$Model %in% diagnoses, ]
ensg1_extra_filtered
```



source file has 1 actual 0 and 92 empty rows(I think?)  
source file has 92 NAs  
```{r}
source_zero <- rna_seq_differential_expression_source[rna_seq_differential_expression_source[[11]] == 0,]
source_na <- rna_seq_differential_expression_source[is.na(rna_seq_differential_expression_source[[11]]),]
print(source_zero)
print(source_na)
```



ENSG00000229807 - has the only row with 0 for pvals in the source data    
- The source data contains one duplicate row for DLPFC AD-CONTROL ALL (Tissue Comparison Sex)    
- The dup rows have different values for the numerics except AveExpr  
- The dup is not the row that has 0 pvals, the zero pvals row is not processed due to Comparison filtering  

```{r}
new_xist <- rna_seq_differential_expression_new[rna_seq_differential_expression_new$ensembl_gene_id == 'ENSG00000229807',]
print(new_xist[order(new_xist$tissue, new_xist$model),])

source_xist <- rna_seq_differential_expression_source[rna_seq_differential_expression_source$ensembl_gene_id == 'ENSG00000229807',]
print(source_xist)

diagnoses <- c("Diagnosis", "Diagnosis.AOD", "Diagnosis.Sex")
diagnoses_filtered_xist <- source_xist[source_xist$Model %in% diagnoses,]
print(diagnoses_filtered_xist)

double_filtered_xist <- diagnoses_filtered_xist[diagnoses_filtered_xist$Comparison == "AD-CONTROL",]

double_filtered_xist[order(double_filtered_xist$Tissue, double_filtered_xist$Model),]
```

Do the 100 new 0s include the 92 NAs? - not entirely overlapping  

-> not sure yet...  


```{r}
source_na_ensg <- source_na$ensembl_gene_id
new_zero <- rna_seq_differential_expression_new[rna_seq_differential_expression_new$adj_p_val == 0,]
new_zero_ensg <- new_zero$ensembl_gene_id


ensgs <- setdiff(new_zero_ensg, source_na_ensg)
cat(sort(ensgs), sep = "\n")

diff <- rna_seq_differential_expression_new[rna_seq_differential_expression_new$ensembl_gene_id %in% ensgs,]
filtered_diff <- diff[diff$adj_p_val == 0,]
print(filtered_diff[order(filtered_diff$ensembl_gene_id, filtered_diff$tissue, filtered_diff$hgnc_symbol),]) # in new not source

reverse_ensgs <- setdiff(source_na_ensg, new_zero_ensg)
cat(sort(reverse_ensgs), sep = "\n")

reverse_diff <-rna_seq_differential_expression_new[rna_seq_differential_expression_new$ensembl_gene_id %in% reverse_ensgs,]
print(reverse_diff[order(reverse_diff$ensembl_gene_id, reverse_diff$tissue, reverse_diff$hgnc_symbol),]) # in source not new
```

## field_length(hgnc_symbol, min = 2, max = 100): 40400 fails

- All of these are due to missing hgnc
- 2342 ENSGs / expect 4 per region for the 7 models so up to 28 dups per ENSG is expected
- but one ENSG has 56
=> dups appear identical, so generally harmless

Resolution: Relax rule

```{r}
v <- validator(field_length(hgnc_symbol, min = 2, max = 100))
cf <- confront(rna_seq_differential_expression_new, v)
out <- values(cf)
is_unique_fails <- rna_seq_differential_expression_new[!out,]
print(is_unique_fails[order(is_unique_fails$ensembl_gene_id, is_unique_fails$tissue, is_unique_fails$model),])
dup_ensg <- as.data.frame(table(is_unique_fails$ensembl_gene_id))
dup_ensg[order(-dup_ensg$Freq),]

```

Rows for the 56 dups case seem identical pairwise, so 28 uniques as expected

```{r}
dups_56 <- rna_seq_differential_expression_new[rna_seq_differential_expression_new$ensembl_gene_id == "ENSG00000229425",]
dups_56[order(dups_56$tissue, dups_56$model),]
```



***
# Other Validation
***

## info for MattC

```{r}
not0 <- rna_seq_differential_expression_new[rna_seq_differential_expression_new$adj_p_val >0,]

max_lfc <- max(rna_seq_differential_expression_new$logfc)
min_lfc <- min(rna_seq_differential_expression_new$logfc)
max_adjP <- max(rna_seq_differential_expression_new$adj_p_val)
min_adjP <- min(not0$adj_p_val)

cat("\nLFC \n  max: ", max_lfc, "\n  min: ", min_lfc, "\nadj_p_val \n  max: ", max_adjP, "\n  min: ", min_adjP)
```
```{r}

clean <- rna_seq_differential_expression_old[!is.na(rna_seq_differential_expression_old$logfc),]

max_lfc <- max(clean$logfc)
min_lfc <- min(clean$logfc)
max_adjP <- max(clean$adj_p_val)
min_adjP <- min(clean$adj_p_val)

cat("\nLFC \n  max: ", max_lfc, "\n  min: ", min_lfc, "\nadj_p_val \n  max: ", max_adjP, "\n  min: ", min_adjP)
```