---
title: "Proteomics coverage analysis"
output: html_notebook
---

```{r eval=FALSE, echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
prepare()
```

# Validation run metadata
```{r echo=FALSE}
# add file-specific values here
name <- "proteomics_diff"
lfq_synId <-  "syn18689335"  
tmt_synId <-  "syn28544692" 
srm_synId <-  "syn25454540" 
```

# Downloaded files
```{r echo=FALSE}
source("~/repos/rstudio/agora-data-validation/src/utils.R")
lfq <- download_file(lfq_synId, type='csv')
tmt <- download_file(tmt_synId, type='csv')
srm <- download_file(srm_synId, type='csv')
```



## proteomic hgnc to ensg correctness (LFQ)

join geneInfo to proteomics on ensembl_gene_id
check if we get the same hgnc_symbol

```{r}
joined <- merge(x = proteomics_new, y = gene_info_new, by = "ensembl_gene_id")
mismatch <- joined[joined$hgnc_symbol.x != joined$hgnc_symbol.y,]
data.frame(mismatch$ensembl_gene_id, mismatch$hgnc_symbol.x, mismatch$hgnc_symbol.y)
```


--------
Coverage
--------

## Identify the names of the columns necessary for comparing the three data sets
```{r}
head(lfq)
head(tmt)
head(srm)
```

LFQ: ENSG, UniqID 
TMT: ENSG, Peptide 
SRM: hgnc_symbol, hgnc_symbol + "|" + Peptide.Name (concatenate cols to emulate identifier) 


## General characterization of the datasets
```{r}
# generate the gene|peptide identifier for srm rows
srm$uid <- paste(srm$hgnc_symbol, "|", srm$UniProtAC, sep = "")

# unique ENSG
lfq_uniq_ensg <- aggregate(lfq$ENSG, by=list(lfq$ENSG), FUN=length)  # 4418
tmt_uniq_ensg <- aggregate(tmt$ENSG, by=list(tmt$ENSG), FUN=length) # 8266

# unique hgnc
lfq_uniq_hgnc <- aggregate(lfq$GeneName, by=list(lfq$GeneName), FUN=length)  # 
tmt_uniq_hgnc <- aggregate(tmt$GeneID, by=list(tmt$GeneID), FUN=length) # 
srm_uniq_hgnc <- aggregate(srm$hgnc_symbol, by=list(srm$hgnc_symbol), FUN=length) # 

# unique gene|peptide identifiers
lfq_uniq <- aggregate(lfq$UniqID, by=list(lfq$UniqID), FUN=length) # 5676 * 4 = 22704 [DLPFC, MFG, TCX, AntPFC]
tmt_uniq <- aggregate(tmt$Peptide, by=list(tmt$Peptide), FUN=length) # 8817
srm_uniq <- aggregate(srm$uid, by=list(srm$uid), FUN=length) # 524

cat("\nLFQ \n\trows:", nrow(lfq), "\n\tunique ENSG", nrow(lfq_uniq_ensg), "\n\tunique hgnc", nrow(lfq_uniq_hgnc), "\n\tunique peptides", nrow(lfq_uniq))
cat("\nTMT \n\trows:", nrow(tmt), "\n\tunique ENSG", nrow(tmt_uniq_ensg), "\n\tunique hgnc", nrow(tmt_uniq_hgnc), "\n\tunique peptides", nrow(tmt_uniq))
cat("\nSRM \n\trows:", nrow(srm), "\n\tunique ENSG", "---", "\n\tunique genes", nrow(srm_uniq_hgnc), "\n\tunique peptides", nrow(srm_uniq))
```



## Overlap
```{r}
install.packages("ggvenn")              
library("ggvenn") 
```

by ENSG (not available for SRM data)
```{r}
ensg <- list('LFQ'=c(lfq$ENSG), 'TMT'=c(tmt$ENSG))
ggvenn(ensg)

```

by hgnc|peptide identifier

```{r}
uniprot_lt <- list('LFQ'=c(lfq$UniqID), 'TMT'=c(tmt$Peptide))

ggvenn(uniprot_lt)
```


by hgnc to include SRM
```{r}
hgnc <- list('LFQ'=c(lfq$GeneName), 'TMT'=c(tmt$GeneID), 'SRM'=c(srm$hgnc_symbol))

ggvenn(hgnc)

```

by hgnc|peptide identifier w/SRM

```{r}
uniprot_lts <- list('LFQ'=c(lfq$UniqID), 'TMT'=c(tmt$Peptide), 'SRM'=c(srm$uid))

ggvenn(uniprot_lts)
```

```{r}
genes <- intersect(srm$hgnc_symbol, tmt$GeneID) # gene names that overlap in the smaller datasets
  
lfq_genes <- lfq[lfq$GeneName %in% genes,]
tmt_genes <- tmt[tmt$GeneID %in% genes,]
srm_genes <- srm[srm$hgnc_symbol%in% genes,]

lfq_genes[order(lfq_genes$GeneName), ]
tmt_genes[order(tmt_genes$GeneID), ]
srm_genes[order(srm_genes$hgnc_symbol), ]
```



```{r}
lfq_uid <- lfq$UniqID
tmt_uid <- tmt$Peptide # 8817 peptides

cat("\nLFQ total rows", length(lfq_uid)) # 22704 peptides
cat("\nTMT total rows", length(tmt_uid)) # 8817 peptides

cat("\nLFQ unique peptides", length(unique(lfq_uid))) # 5676 unique peptides
cat("\nTMT unique peptides", length(unique(tmt_uid))) # 8817 unique peptides

shared <- intersect(lfq_uid, tmt_uid)
lfq_not_tmt <- setdiff(lfq_uid, tmt_uid)
tmt_not_lfq <- setdiff(tmt_uid, lfq_uid)


cat("\nTMT and LFQ", length(shared)) # 3253 genes have both tmt and lfq data
cat("\nLFQ only", length(lfq_not_tmt)) # 2423 genes have tmt data only
cat("\nTMT only", length(tmt_not_lfq)) # 5564 genes have lfq data only
```

```{r}
lfq[lfq$ENSG == 'ENSG00000175416',]
tmt[tmt$ENSG == 'ENSG00000175416',]

lfq[lfq$UniqID == 'CLTB|P09497-2',]
tmt[tmt$Peptide == 'CLTB|P09497-2',]
```


overlap by gene

```{r}
lfq_uniq_ensg <- aggregate(lfq$ENSG, by=list(lfq$ENSG), FUN=length)  # 4418
tmt_uniq_ensg <- aggregate(tmt$ENSG, by=list(tmt$ENSG), FUN=length) # 8266

hist(lfq_uniq_ensg$x)
hist(tmt_uniq_ensg$x)
```